---
title: "Gmacs Example Stock Assessment"
author: "The Gmacs development team"
date: "June 2015"
output:
  pdf_document:
    highlight: zenburn
    toc: yes
  html_document:
    theme: flatly
    toc: yes
  word_document: default
bibliography: Gmacs.bib
---

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.width = 12, fig.height = 8, echo = FALSE, warning = FALSE, message = FALSE)
# Load gmr package for Gmacs:
library(gmr)
```

```{r, load_packages, include=FALSE}

# The model specs
.MODELDIR = c("")
.THEME    = theme_bw(base_size = 12, base_family = "")
.OVERLAY  = TRUE
.SEX      = c("Aggregate","Male","Female")
.FLEET    = c("Pot","Trawl bycatch","NMFS Trawl","BSFRF")
.TYPE     = c("Retained & Discarded","Retained","Discarded")
.SHELL    = c("Aggregate","New Shell","Old Shell")
.MATURITY = c("Aggregate","Immature","Mature")
.SEAS     = c("Annual")

# Read report file and create gmacs report object (a list):
fn       <- paste0(.MODELDIR, "gmacs")
M        <- lapply(fn, read_admb)
names(M) <- c("OneSex")

```

# Model figures

## Fit to survey abundance indices

Gmacs (Figure \ref{fig:survey_biomass}).

```{r survey_biomass, fig.cap = "Model fits to NMFS trawl survey biomass.\\label{fig:survey_biomass}"}
plot_cpue(M, "NMFS Trawl", ylab = "Survey biomass")

```

## Estimated retained catch and discards

The observed and predicted catches by gear type are summarized in (Figure
\ref{fig:fit_to_catch}). 

```{r fit_to_catch, fig.cap = "Observed and predicted catch by gear type for the Gmacs models.\\label{fig:fit_to_catch}"}
# Create a list that only contains the OneSex and TwoSex models
A <- M; A[[3]] <- NULL
plot_catch(A)
```

## Fit to size composition data

The fit of the Gmacs models to the BBRKC size composition data are given in the
following plots. These include fits to the directed pot fishery for males
(Figure \ref{fig:sc_pot_m}), male crabs discarded in the directed pot fishery
(Figure \ref{fig:sc_pot_discarded_m}), female crabs discarded in the directed
pot fishery (Figure \ref{fig:sc_pot_discarded_f}), the groundfish trawl bycatch
fisheries for males (Figure \ref{fig:sc_trawl_bycatch_m}) and females (Figure
\ref{fig:sc_trawl_bycatch_f}), and the NMFS trawl survey (Figure
\ref{fig:sc_NMFS_m}).

```{r sc_pot_m, fig.cap = "Observed and model estimated length-frequencies of male BBRKC by year retained in the directed pot fishery.\\label{fig:sc_pot_m}"}
plot_size_comps(A, 1)
```

```{r sc_pot_discarded_m, fig.cap = "Observed and model estimated length-frequencies of male BBRKC by year discarded in the directed pot fishery.\\label{fig:sc_pot_discarded_m}"}
plot_size_comps(A, 2)
```

```{r sc_pot_discarded_f, fig.cap = "Observed and model estimated length-frequencies of female BBRKC by year discarded in the directed pot fishery.\\label{fig:sc_pot_discarded_f}"}
plot_size_comps(A, 3)
```

```{r sc_trawl_bycatch_m, fig.cap = "Observed and model estimated length-frequencies of male BBRKC by year in the groundfish trawl bycatch fisheries.\\label{fig:sc_trawl_bycatch_m}"}
plot_size_comps(A, 4)
```

```{r sc_trawl_bycatch_f, fig.cap = "Observed and model estimated length-frequencies of female BBRKC by year in the groundfish trawl bycatch fisheries.\\label{fig:sc_trawl_bycatch_f}"}
plot_size_comps(A, 5)
```

```{r , sc_NMFS_m, fig.cap = "Observed and model estimated length-frequencies of male BBRKC by year in the NMFS trawl fishery.\\label{fig:sc_NMFS_m}"}
plot_size_comps(A, 6)
```


## Mean weight-at-length

The mean weight-at-length ($w_\ell$) of crabs is defined in kg and the carapace
length ($\ell$, CL) in mm. The mean weight-at-length of males used in all models
is nearly identical. The only difference between the Gmacs models and Zheng's is
in the final length class (160mm) where the mean weight is greater in Zheng's
model than in Gmacs (Figure \ref{fig:length-weight}). However, the pattern is
very different for females. This difference is due to...

```{r length_weight, fig.cap = "Relationship between carapace length (mm) and weight (kg) by sex in each of the models.\\label{fig:length-weight}"}
.OVERLAY <- FALSE
plot_length_weight(M, xlab = "Carapace length (mm)", ylab = "Weight (kg)")
.OVERLAY <- TRUE
```


## Initial recruitment size distribution

Gmacs was configured to match the @zheng_bristol_2014 model recruitment size
distribution closely (Figure \ref{fig:init_rec}).

```{r init_rec, fig.cap = "Distribution of carapace length (mm) at recruitment.\\label{fig:init_rec}"}
plot_recruitment_size(M, xlab = "Carapace length (mm)")
```


## Molting increment and probability

Options to fit relationship based on data was developed but for the BBRKC
system, a size-specific vector was used to determine molt increments as shown
below (Figure \ref{fig:growth_inc}). Fixed parameters in gmacs were set to
represent that assumed from @zheng_bristol_2014 (Figure \ref{fig:molt_prob}).

```{r growth_inc, fig.cap = "Growth increment (mm).\\label{fig:growth_inc}"}
plot_growth_inc(A)
#plot_growth_inc(M)
```

```{r molt_prob, fig.cap = "Molting probability.\\label{fig:molt_prob}"}
plot_molt_prob(M, xlab = "Carapace length (mm)")
```


## Transition processes

The first set of figures is the growth probabilities (for all crabs that molt)
(Figure \ref{fig:growth_trans}). The second set of figures is the combination of
growth and molting and represents the size transition (Figure
\ref{fig:size_trans}).

```{r growth_trans, fig.cap = "Growth transitions.\\label{fig:growth_trans}"}
plot_growth_transition(M)
```

```{r size_trans, fig.cap = "Size transitions.\\label{fig:size_trans}"}
plot_size_transition(M)
```


## Numbers at length in 1975 and 2014

The number of crabs in each size class (${\bf n}$) in the initial year ($t=1$)
and final year ($t=T$) in each model differ substantially (Figure
\ref{fig:init_N}). The scale of these results differ significantly and may be
related to the interaction with natural mortality estimates and how the initial
population-at-lengths were established (the BBRKC model assumes all new-shell).

```{r init_N, fig.cap = "Numbers at length in 1975.\\label{fig:init_N}"}
plot_numbers(M, c("1975","2014"))
```


## Selectivity

The selectivity by length ($S_\ell$) for each of the fisheries (Figure
\ref{fig:selectivity}).

```{r selectivity, fig.cap = "Estimated selectivity functions.\\label{fig:selectivity}"}
plot_selectivity(A)
```


## Natural mortality

The figure below illustrates implementation of four step changes in $M_t$
(freely estimated) in gmacs relative to the estimates from Zheng et al. 2014
(Figure \ref{fig:M_t}). In both the ADFG-BBRKC and Gmacs-BBRKC models,
time-varying natural mortality ($M_t$) is freely estimated with four step
changes through time. The years ($t$) that each of these steps cover are fixed a
priori. The pattern in time-varying natural mortality is resonably similar
between the two models (Figure \ref{fig:M_t}), however the peak in natural
mortality during the early 1980 is not as high in the Gmacs-BBRKC model.

```{r natural_mortality, fig.cap = "Time-varying natural mortality ($M_t$).\\label{fig:M_t}"}
plot_natural_mortality(M)
```


## Recruitment

Recruitment patterns are similar, but differences in natural mortality
schedules will affect these matches. The figure below plots the values to have
the same mean (Figure \ref{fig:recruitment}). Patterns in recruitment through
time ($R_t$) estimated in the two models are similar, but differences in
natural mortality schedules will affect these matches (Figure
\ref{fig:recruitment}).

```{r recruitment, fig.cap = "Estimated recruitment time series ($R_t$).\\label{fig:recruitment}"}
plot_recruitment(M)
```


## Mature male biomass (MMB)

The spawning stock biomass of mature males, termed the mature male biomass
($\mathit{MMB}_t$), also differs a lot bewteen the two models (Figure
\ref{fig:ssb}).

```{r spawning_stock_biomass, fig.cap = "Mature male biomass (MMB) predicted in the two versions of the Gmacs model (OneSex and TwoSex) and the Zheng model.\\label{fig:ssb}"}
plot_ssb(M)
```


