---
title: "Bristol Bay Red King Crab Stock Assessment 2017"
author: |
  | D'Arcy Webber$^1$, Jie Zheng$^2$, and James Ianelli$^3$
  | $^1$Quantifish, darcy@quantifish.co.nz
  | $^2$Alaska Department of Fish and Game, jie.zheng@alaska.gov
  | $^3$NOAA, jim.ianelli@noaa.gov
date: "`r format(Sys.time(), '%B %Y')`"
output:
  pdf_document:
    includes:
    highlight: zenburn
    keep_tex: yes
    latex_engine: xelatex
  html_document:
    theme: flatly
    toc: yes
  word_document: default
bibliography: ../references/Gmacs.bib
---

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.width = 12, fig.height = 7, echo = FALSE, warning = FALSE, message = FALSE)
```

```{r, load_packages, include=FALSE}
library(gmr)
library(xtable)
options(xtable.comment = FALSE)

# The model specs
mod_names <- c("Gmacs","2016 Model")
#.MODELDIR = c("../../examples/bbrkc_2017/model_1/", "../../examples/bbrkc_2017/model_1/", "../../examples/bbrkc_2017/model_2/","../../examples/bbrkc_2017/model_3/")
.MODELDIR = c("../../examples/bbrkc_2017/model_7/", "../../examples/bbrkc_2017/model_7/")
.THEME    = ggplot2::theme_bw(base_size = 12, base_family = "")
.OVERLAY  = TRUE
.SEX      = c("Aggregate","Male","Female")
.FLEET    = c("Pot","Trawl bycatch","TC bycatch","Fixed Bycatch","NMFS Trawl","BSFRF")
.TYPE     = c("Retained & Discarded","Retained","Discarded")
.SHELL    = c("Aggregate","New Shell","Old Shell")
.MATURITY = c("Aggregate")
.MATURITY = c("Aggregate","Immature","Mature")
.SEAS     = c("1","2","3","4")

# Read report file and create gmacs report object (a list):
fn       <- paste0(.MODELDIR, "gmacs")
M        <- lapply(fn, read_admb)
names(M) <- mod_names

jj <- 2 # The position in the list that Jies model outputs sit
fn <- "2016_model/rk75162"
#fn <- "Jie_2017/rk75162b"
JJ <- lapply(fn, read_admb)

nmult_1 <- 1e+06
nmult_2 <- 0.0004535923 * 1e+6

# Add numbers at length data
# M[[jj]]$N_len[,1] <- nmult_1 * c(JJ[[jj]]$N1, NA)
# M[[jj]]$N_len[,2] <- nmult_1 * c(JJ[[jj]]$N2, NA)
# M[[jj]]$N_len[,3] <- nmult_1 * c(JJ[[jj]]$N3, NA)

# Add MMB data - I think Jie has recorded this in millions of pounds so need to convert to pounds then to tons
# ii <- which(M[[jj]]$fit$names %in% "sd_log_ssb")
# M[[jj]]$ssb <- nmult_2 * JJ[[jj]]$MMB215
# M[[jj]]$fit$est[ii] <- c(log(M[[jj]]$ssb),NA) 
# M[[jj]]$fit$std[ii] <- NA

# Add natural mortality data
#ii <- which(M[[jj]]$mod_yrs == 1998)
#M[[jj]]$M[ii,] <- JJ[[jj]]$M98

# Add recruitment data - It looks like Jie has recorded this as millions of individuals
#din <- JJ[[jj]]$rec
#ii <- which(M[[jj]]$fit$names %in% "sd_log_recruits")
#M[[jj]]$fit$est[ii] <- c(log(nmult_1 * c(din, NA)), NA)
#M[[jj]]$fit$std[ii] <- NA

# Add estimated trawl survey biomass (million of lbs)
# M[[jj]]$pre_cpue[1,] <- c(nmult_2 * JJ[[jj]]$B_ts, NA)
# M[[jj]]$obs_cpue[1,] <- c(nmult_2 * JJ[[jj]]$b_ts, NA)
# M[[jj]]$pre_cpue[2,1:9] <- c(1000 * JJ[[jj]]$X_ps, NA)
# M[[jj]]$obs_cpue[2,1:9] <- c(1000 * JJ[[jj]]$x_ps, NA)
# ii <- which(M[[jj]]$dSurveyData[,1] == 2016)
# M[[jj]]$dSurveyData <- M[[jj]]$dSurveyData[-ii,]
# M[[jj]]$cpue_cv_add[1,39] <- NA
# M[[jj]]$cpue_cv_add[2,9] <- NA
# M[[jj]]$res_cpue[1,39] <- NA
# M[[jj]]$res_cpue[2,9] <- NA

# Add estimated length compositions
#M[[jj]]$d3_pre_size_comps_out[1:15,] <- rbind(t(JJ[[jj]]$P_ob), c(NA,NA,NA))
#M[[jj]]$d3_pre_size_comps_out[16:54,] <- rbind(t(JJ[[jj]]$P_ts), c(NA,NA,NA))
#M[[jj]]$d3_pre_size_comps_out[55:63,] <- rbind(t(JJ[[jj]]$P_ps), c(NA,NA,NA))

# Add selectivity data
# year, sex, fleet, vec
# ind <- which(M[[jj]]$slx_capture[,3] %in% 1)[1:31]
# for (i in ind)
#     M[[jj]]$slx_capture[i,4:6] <- c(JJ[[jj]]$sel_pf[1:2],1) # Directed pot fisheries
# ind <- which(M[[jj]]$slx_capture[,3] %in% 1)[32:38]
# for (i in ind)
#     M[[jj]]$slx_capture[i,4:6] <- c(JJ[[jj]]$sel_pf[3:4],1) # Directed pot fisheries
# ind <- which(M[[jj]]$slx_capture[,3] %in% 4)
# for (i in ind)
#     M[[jj]]$slx_capture[i,4:6] <- c(JJ[[jj]]$sel_ts,1) # Trawl survey
# ind <- which(M[[jj]]$slx_capture[,3] %in% 5)
# for (i in ind)
#     M[[jj]]$slx_capture[i,4:6] <- c(JJ[[jj]]$sel_ps,1) # ADF&G pot survey

# Add recruitment size distribution data - in Jies model all recruit to first size class, in gmacs almost all to first size class.
#M[[jj]]$rec_sdd <- c(1,0,0)

# The .rep files for each of the Gmacs models. Used for making tables of the likelihood components
#like <- list()
#like[[1]] <- readLines("../../examples/bbrkc/OneSex/gmacs.rep")
#like[[2]] <- readLines("../../examples/bbrkc/TwoSex/gmacs.rep")

fn <- paste0(.MODELDIR[2], "gmacs")
Mmatch <- lapply(fn, read_admb)
names(Mmatch) <- c("BBRKC")

fn <- paste0(.MODELDIR[2], "gmacs")
Mbase <- lapply(fn, read_admb)
names(Mbase) <- c("BBRKC")

rinline <- function(code){
  html <- '<code  class="r">``` `CODE` ```</code>'
  sub("CODE", code, html)
}

rec_mod <- 2 # base
#rec_mod <- 4 # M
#rec_mod <- 6 # force
```

# Executive Summary

1. **Stock**: Red king crab (RKC), *Paralithodes camtschaticus*, in Bristol Bay, Alaska.

2. **Catches**: 
The domestic RKC fishery began to expand in the late 1960s and peaked in 1980
with a catch of 129.95 million lbs (58,943 t). The catch declined dramatically in the early
1980s and remained at low levels during the last three decades. Catches during recent years
until 2010/11 were among the high catches in last 15 years. The retained catch in 2015/16
was about 10 million lbs (4,500 t), similar to the catch in 2014/15. The magnitude of
bycatch from groundfish trawl fisheries has been stable and small relative to stock
abundance during the last 10 years.

3. **Stock biomass**: 
Estimated mature biomass increased dramatically in the mid 1970s and
decreased precipitously in the early 1980s. Estimated mature crab abundance had increased
during 1985-2009 with mature females being about three times more abundant in 2009 than
in 1985 and mature males being about two times more abundant in 2009 than in 1985.
Estimated mature abundance has steadily declined since 2009.

4. **Recruitment**: 
Estimated recruitment was high during 1970s and early 1980s and has
generally been low since 1985 (1979 year class). During 1984-2016, only in 1984, 1986,
1995, 1999, 2002 and 2005 were estimated recruitments above the historical average for
1976-2016. Estimated recruitment was extremely low during the last 10 years.

5. **Management performance**: 
Status and catch specifications (1,000 t) (scenario 2) are given below.
In recent assessments, estimated total male catch has been determined as the sum of fishery-reported 
retained catch, estimated male discard mortality in the directed fishery, and estimated male bycatch 
mortality in the groundfish fisheries, as these have been the only sources of non-negligible fishing 
mortality to consider. The stock was above the minimum stock-size threshold (MSST) in 2016/17 and is 
hence not overfished. Overfishing did not occur in 2016/17 (Tables \ref{tab:status} and \ref{tab:status_pounds}).
  
\begin{table}[ht]
\centering
\caption{Status and catch specifications (1000 tons) (scenario {\bf Gmacs base}). Notes: A - calculated from the assessment reviewed by the Crab Plan Team in September 2013, B - calculated from the assessment reviewed by the Crab Plan Team in September 2014, C - calculated from the assessment reviewed by the Crab Plan Team in September 2015, D - calculated from the assessment reviewed by the Crab Plan Team in September 2016.} 
\label{tab:status}
\begin{tabular}{lccccccc}
\hline
          &          & Biomass                        &     & Retained & Total &     & \\ 
Year      & MSST     & ($\mathit{MMB}_\text{mating}$) & TAC & catch    & catch & OFL & ABC \\ 
\hline
2012/13   & 13.19$^A$   & 29.05$^A$   & 3.56 & 3.62 & 3.9  & 7.96 & 7.17 \\
2013/14   & 12.85$^B$   & 27.12$^B$   & 3.90 & 3.99 & 4.56 & 7.07 & 6.36 \\
2014/15   & 13.03$^C$   & 27.25$^C$   & 4.49 & 4.54 & 5.44 & 6.82 & 6.14 \\
2015/16   & 12.89$^D$   & 27.68$^D$   & 4.52 & 4.61 & 5.34 & 6.73 & 6.06 \\
2016/17   &             & 24.00$^D$   &      &      &      & 6.64 & 5.97 \\
\hline
\end{tabular}
\end{table}

\begin{table}[ht]
\centering
\caption{Status and catch specifications (million pounds) (scenario {\bf Gmacs base}).} 
\label{tab:status_pounds}
\begin{tabular}{lcccccccc}
\hline
          &          & Biomass                        &     & Retained & Total &     &     \\ 
Year      & MSST     & ($\mathit{MMB}_\text{mating}$) & TAC & catch    & catch & OFL & ABC \\ 
\hline
2012/13 & 29.1$^A$ & 64.0$^A$         & 7.85 & 7.98           & 8.59        & 17.55 & 15.8  \\
2013/14 & 28.3$^B$ & 59.9$^B$         & 8.6  & 8.8            & 10.05       & 15.58 & 14.02 \\
2014/15 & 28.7$^C$ & 60.1$^C$         & 9.99 & 10.01          & 11.99       & 15.04 & 13.53 \\
2015/16 & 28.4$^D$ & 61.0$^D$         & 9.97 & 10.17          & 11.77       & 14.84 & 13.36 \\
2016/17 &          & 52.9$^D$         &      &                &             & 14.63 & 13.17 \\
\hline
\end{tabular} \\
     
Notes: \\
A – Calculated from the assessment reviewed by the Crab Plan Team in September 2013 \\
B – Calculated from the assessment reviewed by the Crab Plan Team in September 2014 \\
C – Calculated from the assessment reviewed by the Crab Plan Team in September 2015\\
D – Calculated from the assessment reviewed by the Crab Plan Team in September 2016\\
\end{table}


6. **Basis for the OFL**: Estimated mature-male biomass (MMB) on 15 February is used as the measure of biomass for this Tier 4 stock, with males measuring 105 mm CL or more considered mature. The $B_\mathit{MSY}$ proxy is obtained by averaging estimated MMB over a specific reference time period, and current CPT/SSC guidance recommends using the full assessment time frame as the default reference period (Table \ref{tab:ofl_basis}).
\begin{table}[ht]
\centering
\caption{Basis for the OFL (1000 tons) (scenario {\bf Gmacs base}).} 
\label{tab:ofl_basis}
\begin{tabular}{lcccccccc}
\hline
     &      &                  & Biomass                        &                    &                  &          &                            & Natural  \\ 
Year & Tier & $B_\mathit{MSY}$ & ($\mathit{MMB}_\text{mating}$) & $B/B_\mathit{MSY}$ & $F_\mathit{OFL}$ & $\gamma$ & Basis for $B_\mathit{MSY}$ & mortality \\ 
\hline
2012/13 & 3b & 27.5 & 26.3    & 0.96         & 0.31 & 1.0 & 1984-2012 &          0.18 \\
2013/14 & 3b & 26.4 & 25.0    & 0.95         & 0.27 & 1.0 & 1984-2013 &          0.18 \\
2014/15 & 3b & 25.7 & 24.7    & 0.96         & 0.28 & 1.0 & 1984-2014 &          0.18 \\
2015/16 & 3b & 26.1 & 24.7    & 0.95         & 0.27 & 1.0 & 1984-2015 &          0.18 \\
2016/17 & 3b & 25.8 & 24.0    & 0.93         & 0.27 & 1.0 & 1984-2016 &          0.18 \\
\hline
\end{tabular}
\end{table}

\begin{table}[ht]
\centering
\caption{Basis for the OFL (millions of lbs ) (scenario {\bf Gmacs base}).} 
\label{tab:ofl_basis_pounds}
\begin{tabular}{lcccccccc}
\hline
     &      &                  & Biomass                        &                    &                  &          &                            & Natural  \\ 
Year & Tier & $B_\mathit{MSY}$ & ($\mathit{MMB}_\text{mating}$) & $B/B_\mathit{MSY}$ & $F_\mathit{OFL}$ & $\gamma$ & Basis for $B_\mathit{MSY}$ & mortality \\ 
\hline
2012/13 & 3b & 60.7 & 58.0   & 0.96         & 0.31 & 1.0 & 1984-2012 &            0.18 \\
2013/14 & 3b & 58.2 & 55.0   & 0.95         & 0.27 & 1.0 & 1984-2013 &            0.18 \\
2014/15 & 3b & 56.7 & 54.4   & 0.96         & 0.28 & 1.0 & 1984-2014 &            0.18 \\
2015/16 & 3b & 57.5 & 54.4   & 0.95         & 0.27 & 1.0 & 1984-2015 &            0.18 \\
2016/17 & 3b & 56.8 & 52.9   & 0.93         & 0.27 & 1.0 & 1984-2016 &            0.18 \\
\hline
\end{tabular}
\end{table}


# A. Summary of Major Changes

## Changes in Management of the Fishery

There were no new changes in management of the fishery.

## Changes to the Input Data
  a. The new 2016 NMFS trawl survey data and BSFRF side-by-side trawl survey data during
2013-2016 were used.     
  b. Catch and biomass data were updated to include the 2016/17 information.     

## Changes in Assessment Methodology

This assessment was done using Gmacs. 
There are several differences between the
Gmacs assessment model and the previous model. One of the major differences
being that natural and fishing mortality are continuous within `r M[[2]]$nseason` discrete seasons. 
Season length in Gmacs is controlled by
changing the proportion of natural mortality that is applied during each
season. A detailed outline of the Gmacs implementation of the BBRKC model is
provided in Appendix A.

## Changes in Assessment Results

  
# B. Responses to SSC and CPT Comments

## CPT and SSC Comments on Assessments in General

Comment: 
*To come*

Response: 


## CPT and SSC Comments Specific to the BBRKC Stock Assessment

Comment: * [to come ] The SSC and CPT requested the following models for review at the spring 2016 meeting:*

  1. *Base: try to match 2015 model but prevent dome shaped selectivity*
  2. *Base + add CV for both surveys*
  3. *Above + Francis re-weighting*
  4. *Above + remove M spike*

Response:  [to come ] Models 1, 3, and 4 are all included and evaluated in this document as the **Gmacs base**, **Gmacs Francis**, and **Gmacs M** scenarios. Model 2 was not included in this document for two reasons. Firstly, if doing Francis iterative re-weighting then additional CV should not be added as well (as the two methods basically do the same thing). Secondly, the SSC recommended against the model runs with additional CV (see the comment from the SSC below).

Comment: * [to come ] The SSC is not convinced that the model runs with extra CV are very informative. The inclusion of extra CV seems to be rather arbitrary based on the numbers of points that fall within confidence intervals estimated from trawl surveys. The SSC recommends coming up with some alternative way to consider extra variability, which could be informed by simulation testing.*

Response:  [to come ] All model runs that estimate additional CV were dropped from this document. Instead we provide two model runs that use the Francis iterative re-weighting method to re-weight the length-frequency data relative to the abundance indices. These runs are the **Gmacs Francis**, and **Gmacs force** scenarios. The final Gmacs scenario (**Gmacs force**) is an exploratory model run that upweights both the trawl-survey and pot survey abundance indices (it upweights the pot survey more than the trawl survey).


# C. Introduction

## Scientific Name

Red king crab (RKC), *Paralithodes camtschaticus*, in Bristol Bay, Alaska.

## Distribution
Red king crab inhabit intertidal waters to depths >200 m of the North Pacific Ocean from British
Columbia, Canada, to the Bering Sea, and south to Hokkaido, Japan, and are found in several
areas of the Aleutian Islands, eastern Bering Sea, and the Gulf of Alaska.

## Stock Structure
The State of Alaska divides the Aleutian Islands and eastern Bering Sea into three management
registration areas to manage RKC fisheries: Aleutian Islands, Bristol Bay, and Bering Sea
(Alaska Department of Fish and Game (ADF&G) 2012). The Bristol Bay area includes all waters
north of the latitude of Cape Sarichef (54°36' N lat.), east of 168°00' W long., and south of the
latitude of Cape Newenham (58°39' N lat.) and the fishery for RKC in this area is managed
separately from fisheries for RKC outside of this area; i.e., the red king crab in the Bristol Bay
area are assumed to be a separate stock from red king crab outside of this area. This report
summarizes the stock assessment results for the Bristol Bay RKC stock.

## Life History
Red king crab have a complex life history. Fecundity is a function of female size, ranging from
several tens of thousands to a few hundreds of thousands (Haynes 1968; Swiney et al. 2012). The
eggs are extruded by females, fertilized in the spring, and held by females for about 11 months
(Powell and Nickerson 1965). Fertilized eggs are hatched in the spring, most during April-June
(Weber 1967). Primiparous females are bred a few weeks earlier in the season than multiparous
females.
Larval duration and juvenile crab growth depend on temperature (Stevens 1990; Stevens and
Swiney 2007). Male and female RKC mature at 5–12 years old, depending on stock and
temperature (Loher et al. 2001; Stevens 1990) and may live >20 years (Matsuura and Takeshita
1990). Males and females attain a maximum size of 227 and 195 mm carapace length (CL),
respectively (Powell and Nickerson 1965). Female maturity is evaluated by the size at which
females are observed to carry egg clutches. Male maturity can be defined by multiple criteria
including spermataphore production and size, chelae vs. carapace allometry, and participation in
mating in situ (reviewed by Webb 2014). For management purposes, females >89 mm CL and
males >119 mm CL are assumed to be mature for Bristol Bay RKC. Juvenile RKC molt multiple
times per year until age 3 or 4; thereafter, molting continues annually in females for life and in
males until maturity. Male molting frequency declines after attaining functional maturity.

## Fishery
The RKC stock in Bristol Bay, Alaska, supports one of the most valuable fisheries in the United
States. A review of the history of the Bristol Bay RKC fishery is provided in Fitch et al. (2012) and
Otto (1989). The Japanese fleet started the fishery in the early 1930s, stopped fishing from 1940 to
1952, and resumed the fishery from 1953 until 1974. The Russian fleet fished for RKC from 1959
to 1971. The Japanese fleet employed primarily tanglenets with a very small proportion of catch
from trawls and pots. The Russian fleet used only tanglenets. United States trawlers started fishing
Bristol Bay RKC in 1947, but the effort and catch declined in the 1950s. The domestic RKC fishery
began to expand in the late 1960s and peaked in 1980 with a catch of 129.95 million lbs (58,943 t),
worth an estimated $115.3 million ex-vessel value. The catch declined dramatically in the early
1980s and has remained at low levels during the last two decades (Table 1). After the early 1980s
stock collapse, the Bristol Bay RKC fishery took place during a short period in the fall (usually
lasting about a week) with the catch quota based on the stock assessment conducted the previous
summer (Zheng and Kruse 2002). Beginning with the 2005/2006 season, new regulations associated
with fishery rationalization resulted in an increase in the duration of the fishing season (October 15
to January 15). With the implementation of crab rationalization, historical guideline harvest levels
(GHL) were changed to a total allowable catch (TAC). Before rationalization, the implementation
errors were quite high for some years and total actual catch from 1980 to 2007 was about 6% less
than the sum of GHL/TAC over that period.

## Fisheries Management
King and Tanner crab stocks in the Bering Sea and Aleutian Islands are managed by the State of
Alaska through a federal king and Tanner crab fishery management plan (FMP). Under the FMP,
management measures are divided into three categories: (1) fixed in the FMP, (2) frame worked in
the FMP, and (3) discretion of the State of Alaska. The State of Alaska is responsible for
determining and establishing the GHL/TAC under the framework in the FMP.
Harvest strategies for the Bristol Bay RKC fishery have changed over time. Two major
management objectives for the fishery are to maintain a healthy stock that ensures reproductive
viability and to provide for sustained levels of harvest over the long term (ADF&G 2012). In
attempting to meet these objectives, the GHL/TAC is coupled with size-sex-season restrictions.
Only males ≥6.5-in carapace width (equivalent to 135-mm carapace length, CL) may be
harvested and no fishing is allowed during molting and mating periods (ADF&G 2012).
Specification of TAC is based on a harvest rate strategy. Before 1990, harvest rates on legal
males were based on population size, abundance of prerecruits to the fishery, and postrecruit
abundance, and rates varied from less than 20% to 60% (Schmidt and Pengilly 1990). In 1990,
the harvest strategy was modified, and a 20% mature male harvest rate was applied to the
abundance of mature-sized (≥120-mm CL) males with a maximum 60% harvest rate cap of legal
(≥135-mm CL) males (Pengilly and Schmidt 1995). In addition, a minimum threshold of 8.4
million mature-sized females (≥90-mm CL) was added to existing management measures to
avoid recruitment overfishing (Pengilly and Schmidt 1995). Based on a new assessment model
and research findings (Zheng et al. 1995a, 1995b, 1997a, 1997b), the Alaska Board of Fisheries
adopted a new harvest strategy in 1996. That strategy had two mature male harvest rates: 10%
when effective spawning biomass (ESB) is between 14.5 and 55.0 million lbs and 15% when
ESB is at or above 55.0 million lbs (Zheng et al. 1996). The maximum harvest rate cap of legal
males was changed from 60% to 50%. A threshold of 14.5 million lbs of ESB was also added. In
1997, a minimum threshold of 4.0 million lbs was established as the minimum GHL for opening
the fishery and maintaining fishery manageability when the stock abundance is low. The Board
modified the current harvest strategy by adding a mature harvest rate of 12.5% when the ESB is
between 34.75 and 55.0 million lbs in 2003 and eliminated the minimum GHL threshold in 2012.
The current harvest strategy is illustrated in Figure 1.


\ref{tab:bbrkc_fishery}).

\begin{table}[ht]
\centering
\caption{ Bristol Bay red king crab annual catch and bycatch mortality biomass (t) from
June 1 to May 31. A handling mortality rate of 20% for the directed pot, 25%
for the Tanner fishery, and 80% for trawl was assumed to estimate bycatch
mortality biomass. }
\label{tab:bbrkc_fishery}
\footnotesize
\begin{tabular}{lrrrrrrrrr}
\hline
Year & \multicolumn{4}{c}{Retained Catch}                 & \multicolumn{2}{c}{Pot Bycatch}Trawl         & Tanner Fshry  & Total      \\
     & U.S.           & Cost-Recovery & Foreign & Total   & Males       & Females & Bycatch       & Bycatch        & Catch      \\
\hline
1953 & 1331.3         &               & 4705.6  & 6036.9  &             &         &               &                & 6036.9     \\
1954 & 1149.9         &               & 3720.4  & 4870.2  &             &         &               &                & 4870.2      \\
1955 & 1029.2         &               & 3712.7  & 4741.9  &             &         &               &                & 4741.9      \\
1956 & 973.4          &               & 3572.9  & 4546.4  &             &         &               &                & 4546.4      \\
1957 & 339.7          &               & 3718.1  & 4057.8  &             &         &               &                & 4057.8      \\
1958 & 3.2            &               & 3541.6  & 3544.8  &             &         &               &                & 3544.8      \\
1959 & 0              &               & 6062.3  & 6062.3  &             &         &               &                & 6062.3      \\
1960 & 272.2          &               & 12200.7 & 12472.9 &             &         &               &                & 12472.9     \\
1961 & 193.7          &               & 20226.6 & 20420.3 &             &         &               &                & 20420.3     \\
1962 & 30.8           &               & 24618.7 & 24649.6 &             &         &               &                & 24649.6     \\
1963 & 296.2          &               & 24930.8 & 25227   &             &         &               &                & 25227       \\
1964 & 373.3          &               & 26385.5 & 26758.8 &             &         &               &                & 26758.8     \\
1965 & 648.2          &               & 18730.6 & 19378.8 &             &         &               &                & 19378.8     \\
1966 & 452.2          &               & 19212.4 & 19664.6 &             &         &               &                & 19664.6     \\
1967 & 1407           &               & 15257   & 16664.1 &             &         &               &                & 16664.1     \\
1968 & 3939.9         &               & 12459.7 & 16399.6 &             &         &               &                & 16399.6     \\
1969 & 4718.7         &               & 6524    & 11242.7 &             &         &               &                & 11242.7     \\
1970 & 3882.3         &               & 5889.4  & 9771.7  &             &         &               &                & 9771.7      \\
1971 & 5872.2         &               & 2782.3  & 8654.5  &             &         &               &                & 8654.5      \\
1972 & 9863.4         &               & 2141    & 12004.3 &             &         &               &                & 12004.3     \\
1973 & 12207.8        &               & 103.4   & 12311.2 &             &         &               &                & 12311.2     \\
1974 & 19171.7        &               & 215.9   & 19387.6 &             &         &               &                & 19387.6     \\
1975 & 23281.2        &               & 0       & 23281.2 &             &         &               &                & 23281.2     \\
1976 & 28993.6        &               & 0       & 28993.6 &             &         & 682.8         &                & 29676.4     \\
1977 & 31736.9        &               & 0       & 31736.9 &             &         & 1249.9        &                & 32986.8     \\
1978 & 39743          &               & 0       & 39743   &             &         & 1320.6        &                & 41063.6     \\
1979 & 48910          &               & 0       & 48910   &             &         & 1331.9        &                & 50241.9     \\
1980 & 58943.6        &               & 0       & 58943.6 &             &         & 1036.5        &                & 59980.1     \\
1981 & 15236.8        &               & 0       & 15236.8 &             &         & 219.4         &                & 15456.2     \\
1982 & 1361.3         &               & 0       & 1361.3  &             &         & 574.9         &                & 1936.2      \\
1983 & 0              &               & 0       & 0       &             &         & 420.4         &                & 420.4       \\
1984 & 1897.1         &               & 0       & 1897.1  &             &         & 1094          &                & 2991.1      \\
1985 & 1893.8         &               & 0       & 1893.8  &             &         & 390.1         &                & 2283.8      \\
1986 & 5168.2         &               & 0       & 5168.2  &             &         & 200.6         &                & 5368.8      \\
1987 & 5574.2         &               & 0       & 5574.2  &             &         & 186.4         &                & 5760.7      \\
1988 & 3351.1         &               & 0       & 3351.1  &             &         & 597.8         &                & 3948.9      \\
1989 & 4656           &               & 0       & 4656    &             &         & 174.1         &                & 4830.1      \\
1990 & 9236.2         & 36.6          & 0       & 9272.8  & 526.9       & 651.5   & 247.6         &                & 10698.7     \\
1991 & 7791.8         & 93.4          & 0       & 7885.1  & 407.8       & 75      & 316           & 1401.8         & 10085.7     \\
1992 & 3648.2         & 33.6          & 0       & 3681.8  & 552         & 418.5   & 335.4         & 244.4          & 5232.2      \\
1993 & 6635.4         & 24.1          & 0       & 6659.6  & 763.2       & 637.1   & 426.6         & 54.6           & 8541        \\
1994 & 0              & 42.3          & 0       & 42.3    & 3.8         & 1.9     & 88.9          & 10.8           & 147.8       \\
1995 & 0              & 36.4          & 0       & 36.4    & 3.3         & 1.6     & 194.2         & 0              & 235.5       \\
1996 & 3812.7         & 49            & 0       & 3861.7  & 164.6       & 1       & 106.5         & 0              & 4133.9      \\
1997 & 3971.9         & 70.2          & 0       & 4042.1  & 244.7       & 19.6    & 73.4          & 0              & 4379.8      \\
1998 & 6693.8         & 85.4          & 0       & 6779.2  & 959.7       & 864.9   & 159.8         & 0              & 8763.7      \\
1999 & 5293.5         & 84.3          & 0       & 5377.9  & 314.2       & 8.8     & 201.6         & 0              & 5902.4      \\
2000 & 3698.8         & 39.1          & 0       & 3737.9  & 360.8       & 40.5    & 100.4         & 0              & 4239.5      \\
2001 & 3811.5         & 54.6          & 0       & 3866.2  & 417.9       & 173.5   & 164.6         & 0              & 4622.1      \\
2002 & 4340.9         & 43.6          & 0       & 4384.5  & 442.7       & 7.3     & 155.1         & 0              & 4989.6      \\
2003 & 7120           & 15.3          & 0       & 7135.3  & 918.9       & 430.4   & 172.3         & 0              & 8656.9      \\
2004 & 6915.2         & 91.4          & 0       & 7006.7  & 345.5       & 187     & 119.6         & 0              & 7658.8      \\
2005 & 8305           & 94.7          & 0       & 8399.7  & 1359.5      & 498.3   & 155.2         & 0              & 10412.8     \\
2006 & 7005.3         & 137.9         & 0       & 7143.2  & 563.8       & 37      & 116.7         & 3.8            & 7864.4      \\
2007 & 9237.9         & 66.1          & 0       & 9303.9  & 1001.3      & 186.1   & 138.5         & 1.8            & 10631.6     \\
2008 & 9216.1         & 0             & 0       & 9216.1  & 1165.5      & 148.4   & 159.5         & 4              & 10693.5     \\
2009 & 7226.9         & 45.5          & 0       & 7272.5  & 888.1       & 85.2    & 103.7         & 1.6            & 8351.2      \\
2010 & 6728.5         & 33            & 0       & 6761.5  & 797.5       & 122.6   & 85.3          & 0              & 7767        \\
2011 & 3553.3         & 53.8          & 0       & 3607.1  & 395         & 24      & 68.8          & 0              & 4094.9      \\
2012 & 3560.6         & 61.1          & 0       & 3621.7  & 205.2       & 12.3    & 61.2          & 0              & 3900.5      \\
2013 & 3901.1         & 89.9          & 0       & 3991    & 310.6       & 99.8    & 136.2         & 28.5           & 4566        \\
2014 & 4530           & 8.6           & 0       & 4538.6  & 584.7       & 86.2    & 221.9         & 42             & 5473.4      \\
2015 & 4522.3         & 91.4          & 0       & 4613.7  & 266.1       & 222.9   & 149.4         & 84.2           & 5336.3      \\
\hline
\end{tabular}
\end{table}

\begin{table}[ht]
\centering
\caption{Observed proportion of crab by size class during the ADF\&G crab observer pot-lift sampling. Source: ADF\&G Crab Observer Database.} 
\label{tab:stage_cpue_1}
\begin{tabular}{lrrrrrr}
  \hline
  Year & Total pot lifts & Pot lifts sampled & Number of crab (90 mm+ CL) & Stage 1 & Stage 2 & Stage 3 \\
  \hline
  1990/91 & 26,264 &  10 &   150 & 0.113 & 0.393 & 0.493 \\
  1991/92 & 37,104 & 125 & 3,393 & 0.133 & 0.177 & 0.690 \\
  1992/93 & 56,630 &  71 & 1,606 & 0.191 & 0.268 & 0.542 \\
  1993/94 & 58,647 &  84 & 2,241 & 0.281 & 0.210 & 0.510 \\
  1994/95 & 60,860 & 203 & 4,735 & 0.294 & 0.271 & 0.434 \\
  1995/96 & 48,560 &  47 &   663 & 0.148 & 0.212 & 0.640 \\
  1996/97 & 91,085 &  96 &   489 & 0.160 & 0.223 & 0.618 \\
  1997/98 & 81,117 & 133 & 3,195 & 0.182 & 0.205 & 0.613 \\
  1998/99 & 91,826 & 135 & 1.322 & 0.193 & 0.216 & 0.591 \\
  \multicolumn{2}{l}{1999/00 - 2008/09} & \multicolumn{5}{c}{FISHERY CLOSED} \\
  2009/10 & 10,484 & 989   & 19,802 & 0.141 & 0.324 & 0.535 \\
  2010/11 & 29,356 & 2,419 & 45,466 & 0.131 & 0.315 & 0.553 \\
  2011/12 & 48,554 & 3,359 & 58,666 & 0.131 & 0.305 & 0.564 \\
  2012/13 & 37,065 & 2,841 & 57,298 & 0.141 & 0.318 & 0.541 \\
  \multicolumn{2}{l}{2013/14} & \multicolumn{5}{c}{FISHERY CLOSED} \\
  2014/15 & 10,133 & 895 & 9,906 & 0.094 & 0.228 & 0.679 \\
  2015/16 &  5,475 & 419 & 3,248 & 0.115 & 0.252 & 0.633 \\
  \hline
\end{tabular}
\end{table}

```{r bbrkc_groundfish_bycatch, results = "asis"}
#df <- dplyr::filter(data.frame(M[[2]]$dCatchData_out), V2 == 2, V3 %in% c(2,3)) %>%
#  dplyr::select(V1, V3, V5) %>%
#  tidyr::spread(key = V3, value = V5)
#names(df) <- c("Year","Trawl bycatch","Fixed gear bycatch")
#tab <- xtable(df, caption = "Groundfish BBRKC male bycatch biomass (tons) estimates. Trawl includes pelagic trawl and non-pelagic trawl types. Source: J. Zheng, ADF\\&G, and author estimates based on data from R. Foy, NMFS. AKRO estimates used after 2008/09.", label = "tab:bbrkc_groundfish_bycatch", digits = c(0,0,3,3))
#print(tab, caption.placement = "top", include.rownames = FALSE)
```


# D. Data

## Summary of New Information

Data used in this assessment have been updated to include the most recently
available fishery and survey numbers. This assessment makes use of two new
survey data points including the 2016 NMFS trawl-survey estimate of abundance,
and the 2016 ADF&G pot survey CPUE. Both of these surveys have associated size
compositon data. The assessment also uses updated 1993-2015 groundfish and
fixed gear bycatch estimates based on AKRO data. The 2015/16 directed fishery
catch data and associated size composition data were also used. The data used
in each of the new models is shown in Figure \ref{fig:data_extent}.

```{r data_extent, fig.cap = "Data extent for the BBRKC assessment.\\label{fig:data_extent}"}
plot_datarange(Mbase)
```

## Major Data Sources

Major data sources used in this assessment include annual directed-fishery retained-catch statistics from fish tickets (1978/79-1998/99, 2009/10-2012/13, and 2014/15-2015/16; Table \ref{tab:bbrkc_fishery}); results from the annual NMFS eastern Bering Sea trawl survey (1978-2016; Table \ref{tab:stage_cpue_nmfs}); results from the triennial ADF&G bbrkC pot survey (every third year during 1995-2013), the 2015 pot survey, and the 2016 pot survey (Table \ref{tab:stage_cpue}); size-frequency information from ADF&G crab-observer pot-lift sampling (1990/91-1998/99, 2009/10-2012/13, and 2014/15-2015/16; Table \ref{tab:stage_cpue_1}); and NMFS groundfish-observer bycatch biomass estimates (1992/93-2015/16; Table \ref{tab:smbkc_groundfish_bycatch}). 
\begin{table}[ht]
\centering
\caption{Size-class and total CPUE (90+ mm CL) with estimated CV and total number of captured crab (90+ mm CL) from the 96 common stations surveyed during the seven triennial ADF\&G SMBKC pot surveys and the 2015 and 2016 surveys. Source: D. Pengilly and R. Gish, ADF\&G.} 
\label{tab:stage_cpue}
\begin{tabular}{lcccrrr}
  \hline
       & Stage-1     & Stage-2      & Stage-3   &            &    & \\ 
  Year & (90-104 mm) & (105-119 mm) & (120+ mm) & Total CPUE & CV & Number of crabs \\ 
  \hline
  1995 & 1.919 & 3.198 & 6.922 & 12.042 & 0.13 & 4624 \\ 
  1998 & 0.964 & 2.763 & 8.804 & 12.531 & 0.06 & 4812 \\ 
  2001 & 1.266 & 1.737 & 5.487 & 8.477 & 0.08 & 3255 \\ 
  2004 & 0.112 & 0.414 & 1.141 & 1.667 & 0.15 & 640 \\ 
  2007 & 1.086 & 2.721 & 4.836 & 8.643 & 0.09 & 3319 \\ 
  2010 & 1.326 & 3.276 & 5.607 & 10.209 & 0.13 & 3920 \\ 
  2013 & 0.878 & 1.398 & 3.367 & 5.643 & 0.19 & 2167 \\ 
  2015 & 0.198 & 0.682 & 1.924 & 2.805 & 0.18 & 1077 \\ 
  2016 & 0.198 & 0.456 & 1.724 & 2.378 & 0.186 & 777 \\ 
   \hline
\end{tabular}
\end{table}
\begin{table}[ht]
\centering
\caption{NMFS EBS trawl-survey area-swept estimates of male crab abundance ($10^6$ crab) and of mature male biomass ($10^6$ lbs). Total number of captured male crab $\ge$ 90 mm CL is also given. Source: R. Foy, NMFS. The "+" refer to plus group.} 
\label{tab:stage_cpue_nmfs}
\begin{tabular}{lccccccccc}
  \hline
       & \multicolumn{5}{c}{Abundance}                       & & \multicolumn{2}{c}{Biomass} & \\
  \cline{2-6}\cline{8-9}
       & Stage-1     & Stage-2      & Stage-3   &       &    & & Total       &               & Number \\ 
  Year & (90-104 mm) & (105-119 mm) & (120+ mm) & Total & CV & & (90+ mm CL) & CV            & of crabs \\ 
  \hline
  1978 & 2.213 & 1.991 & 1.521 & 5.726 & 0.411 & & 15.064 & 0.394 & 157 \\
  1979 & 3.061 & 2.281 & 1.808 & 7.150 & 0.472 & & 17.615 & 0.463 & 178 \\
  1980 & 2.856 & 2.563 & 2.541 & 7.959 & 0.572 & & 22.017 & 0.507 & 185 \\
  1981 & 0.483 & 1.213 & 2.263 & 3.960 & 0.368 & & 14.443 & 0.402 & 140 \\
  1982 & 1.669 & 2.431 & 5.884 & 9.984 & 0.401 & & 35.763 & 0.344 & 271 \\
  1983 & 1.061 & 1.651 & 3.345 & 6.057 & 0.332 & & 21.240 & 0.298 & 231 \\
  1984 & 0.435 & 0.497 & 1.452 & 2.383 & 0.175 & & 8.976 & 0.179 & 105 \\
  1985 & 0.379 & 0.376 & 1.117 & 1.872 & 0.216 & & 6.858 & 0.210 & 93 \\
  1986 & 0.203 & 0.447 & 0.374 & 1.025 & 0.428 & & 3.124 & 0.388 & 46 \\
  1987 & 0.325 & 0.631 & 0.715 & 1.671 & 0.302 & & 5.024 & 0.291 & 71 \\
  1988 & 0.410 & 0.816 & 0.957 & 2.183 & 0.285 & & 6.963 & 0.252 & 81 \\
  1989 & 2.169 & 1.154 & 1.786 & 5.109 & 0.314 & & 13.974 & 0.271 & 208 \\
  1990 & 1.053 & 1.031 & 2.338 & 4.422 & 0.302 & & 14.837 & 0.274 & 170 \\
  1991 & 1.147 & 1.665 & 2.233 & 5.046 & 0.259 & & 15.318 & 0.248 & 197 \\
  1992 & 1.074 & 1.382 & 2.291 & 4.746 & 0.206 & & 15.638 & 0.201 & 220 \\
  1993 & 1.521 & 1.828 & 3.276 & 6.626 & 0.185 & & 21.051 & 0.169 & 324 \\
  1994 & 0.883 & 1.298 & 2.257 & 4.438 & 0.187 & & 14.416 & 0.176 & 211 \\
  1995 & 1.025 & 1.188 & 1.741 & 3.953 & 0.187 & & 12.574 & 0.178 & 178 \\
  1996 & 1.238 & 1.891 & 3.064 & 6.193 & 0.263 & & 20.746 & 0.241 & 285 \\
  1997 & 1.165 & 2.228 & 3.789 & 7.182 & 0.367 & & 24.084 & 0.337 & 296 \\
  1998 & 0.660 & 1.661 & 2.849 & 5.170 & 0.373 & & 17.586 & 0.355 & 243 \\
  1998 & 0.223 & 0.222 & 0.558 & 1.003 & 0.192 & & 3.515 & 0.182 & 52 \\
  2000 & 0.282 & 0.285 & 0.740 & 1.307 & 0.303 & & 4.623 & 0.310 & 61 \\
  2001 & 0.419 & 0.502 & 0.938 & 1.859 & 0.243 & & 6.242 & 0.245 & 91 \\
  2002 & 0.111 & 0.230 & 0.640 & 0.981 & 0.311 & & 3.820 & 0.320 & 38 \\
  2003 & 0.449 & 0.280 & 0.465 & 1.194 & 0.399 & & 3.454 & 0.336 & 65 \\
  2004 & 0.247 & 0.184 & 0.562 & 0.993 & 0.369 & & 3.360 & 0.305 & 48 \\
  2005 & 0.319 & 0.310 & 0.501 & 1.130 & 0.403 & & 3.620 & 0.371 & 42 \\
  2006 & 0.917 & 0.642 & 1.240 & 2.798 & 0.339 & & 8.585 & 0.334 & 126 \\
  2007 & 2.518 & 2.020 & 1.193 & 5.730 & 0.420 & & 14.266 & 0.385 & 250 \\
  2008 & 1.352 & 0.801 & 1.457 & 3.609 & 0.289 & & 10.261 & 0.284 & 167 \\
  2009 & 1.573 & 2.161 & 1.410 & 5.144 & 0.263 & & 13.892 & 0.256 & 251 \\
  2010 & 3.937 & 3.253 & 2.458 & 9.648 & 0.544 & & 24.539 & 0.466 & 388 \\
  2011 & 1.800 & 3.255 & 3.207 & 8.263 & 0.587 & & 24.099 & 0.558 & 318 \\
  2012 & 0.705 & 1.970 & 1.808 & 4.483 & 0.361 & & 13.669 & 0.339 & 193 \\
  2013 & 0.335 & 0.452 & 0.807 & 1.593 & 0.215 & & 5.043 & 0.217 & 74 \\
  2014 & 0.723 & 1.627 & 1.809 & 4.160 & 0.503 & & 13.292 & 0.449 & 181 \\
  2015 & 0.992 & 1.269 & 1.979 & 4.240 & 0.774 & & 12.958 & 0.770 & 153 \\
  2016 & 0.535 & 0.660 & 1.178 & 2.373 & 0.447 & & 7.685 & 0.393 & 108 \\
  \hline
\end{tabular}
\end{table}

XXXFigure \ref{fig:stations} maps stations from which SMBKC trawl-survey and pot-
survey data were obtained. Further information concerning the NMFS trawl
survey as it relates to commercial crab species is available in Daly et al.
(2014); see Gish et al. (2012) for a description of ADF&G SMBKC pot-survey
methods. It should be noted that the two surveys cover different geographic
regions and that each has in some years encountered proportionally large
numbers of male blue king crab in areas where the other is not represented
(Figure \ref{fig:catch181}). Crab-observer sampling protocols are detailed in
the crab-observer training manual (ADF&G 2013). Groundfish SMBKC bycatch data
come from NMFS Bering Sea reporting areas 521 and 524 (Figure
\ref{fig:reporting_areas}). Note that for this assessment the newly available
NMFS groundfish observer data reported by ADF&G statistical area was not used.

## Other Data Sources

XXXRecent model configurations developed for SMBKC makes use of a growth
transition matrix based on Otto and Cummiskey (1990), the same growth
transition matrix is used in this assessment. Other relevant data sources,
including assumed population and fishery parameters, are presented in Appendix
A, which also provides a detailed description of the model configuration used
for this assessment.

## Excluded Data Sources

Groundfish bycatch size-frequency data are available for selected years. These data were used in model-based assessments prior to 2011. However, they have since been excluded because these data tend to be severely limited: for example, 2012/13 data include a total of just 4 90 mm+ CL male blue king crab from reporting areas 521 and 524.


\newpage\clearpage

# E. Analytic Approach

## History of Modeling Approaches for this Stock

XXXA four-stage catch-survey-analysis (CSA) assessment model was used before 2011
to estimate abundance and biomass and prescribe fishery quotas for the SMBKC
stock (2010 SAFE; Zheng et al. 1997). The four-stage CSA is similar to a full
length-based analysis, the major difference being coarser length groups, which
are more suited to a small stock with consistently low survey catches. In this
approach, the abundance of male crab with a CL of 90 mm or above is modeled in
terms of four crab stages: stage 1: 90-104 mm CL; stage 2: 105-119 mm CL;
stage 3: newshell 120-133 mm CL; and stage 4: oldshell $\ge$ 120 mm CL and
newshell $\ge$ 134 mm CL. Motivation for these stage definitions comes from
the fact that for management of the SMBKC stock, male crab measuring at least
105 mm CL are considered mature, whereas 120 mm CL is considered a proxy for
the legal size of 5.5 in carapace width, including spines. Additional
motivation for these stage definitions comes from an estimated average growth
increment of about 14 mm per molt for SMBKC (Otto and Cummiskey 1990).

## Assessment Methodology

The 2016 SMBKC assessment model makes use of the modeling framework Gmacs. The aim when developing this model was to first provide a fit to the data that best matched the 2015 SMBKC stock assessment model. A detailed description of the Gmacs model and its implementation is presented in Appendix A.

## Model Selection and Evaluation

Five different Gmacs model scenarios were considered, in this document results from these models and the 2015 model are compared. The models inlcude:

1. **2015 Model**: the 2015 approach with a correction^[A correction to the 2015 model code was made in the population dynamics function involving how the growth transition matrix was applied to the numbers at length to calculate the numbers during the following time-step, specifically `r rinline("N(t+1,3)=TM(2,3)*NN(2)+NN(3);")` was changed to `r rinline("N(t+1,3)=TM(1,3)*NN(1)+TM(2,3)*NN(2)+NN(3);")`.]. This modification was made prior to comparisons (note that this modification caused the NMFS trawl survey selectivity to exceed 1 for stage-2 crab).

2. **Gmacs match**: tries to match as closely as possible with the 2015 Model by fixing the stage-1 and stage-2 selectivity parameters and the catchability coefficient ($q$) for the ADF&G pot survey at those values estimated in the 2015 model (and allows the NMFS trawl survey selectivity to exceed 1 for stage-2 crab). The parameters that are estimated in this model include the average recruitment ($\bar{R}$), the recruitment deviations ($\delta^R_y$), the initial numbers in each stage ($\boldsymbol{n}^0$), the natural mortality deviation 1998 ($\delta^M_{1998}$), and the fishing mortalities for the directed pot fishery, the trawl bycatch fishery, and the fixed bycatch fishery ($\bar{F}^\text{df}$, $\bar{F}^\text{tb}$, $\bar{F}^\text{fb}$, $\delta^\text{df}_{t,y}$, $\delta^\text{tb}_{t,y}$, $\delta^\text{fb}_{t,y}$). As in the 2015 model, the robust multinomial distribution was used to model the length-frequency data.

3. **Gmacs base**: directed pot, NMFS trawl survey and ADF&G pot survey selectivities are estimated for stage-1 and stage-2 crab (and fixed at 1 for stage-3 crab). These selectivities are bounded so that they cannot be greater than 1. This model also estimates the catchability coefficient ($q$) for the ADF&G pot survey as well as the average recruitment ($\bar{R}$), the recruitment deviations ($\delta^R_y$), the initial numbers in each stage ($\boldsymbol{n}^0$), the natural mortality deviation 1998 ($\delta^M_{1998}$), and the fishing mortalities for the directed pot fishery, the trawl bycatch fishery, and the fixed bycatch fishery ($\bar{F}^\text{df}$, $\bar{F}^\text{tb}$, $\bar{F}^\text{fb}$, $\delta^\text{df}_{t,y}$, $\delta^\text{tb}_{t,y}$, $\delta^\text{fb}_{t,y}$). As in the 2015 model, the robust multinomial distribution was used to model the length-frequency data.

4. **Gmacs M**: is the same as above except that natural mortality ($M$) is fixed at 0.18 $\text{yr}^{-1}$ during all years.

5. **Gmacs Francis**: is similar to the scenario above except that it also uses the Francis iterative re-weighting method (Francis 2011), to re-weight the size-composition data relative to the abundance indices. The trawl survey and pot survey weights were left as is (i.e. a weight of 1) because upweighting these series resulted in worse standard deviation of the normalised residual (SDNR) and median of the absolute residual (MAR) values for each of the surveys. Down-weighting the two surveys actually improved the SDNR and MAR values, but it would be unwise to down-weight either of these series. When applying the Francis iterative re-weighting method only once iteration was done (i.e. the model was run once with the size composition likelihood weights set to one, the new Francis weights were calculated, and the model was run once more using these weights). In this scenario the multinomial distribution was used instead as the theory underpinning the Francis weighting method is based on this distribution.

6. **Gmacs force**: is an exploratory scenario that the same as above except the NMFS trawl survey is up-weighted by $\lambda^\text{NMFS}=$ `r M[[2]]$cpue_lambda[1]` and the ADF&G pot survey is up-weighted by $\lambda^\text{ADFG}=$ `r M[[2]]$cpue_lambda[2]`. After this, the Francis weights for each of the size-compostitons were recalculated and applied again in this model. This scenario should not be used for overfishing determination as it upweights the trawl and pot survey abundance indices to force a better fit to each of these data sets and provide some contrast among the Gmacs model runs. This scenario forces a better fit to the trawl and pot surveys at the expense of the SDNR (and MAR) for each of these series.

Table \ref{tab:model_runs} outlines the major features of each of the models.
\begin{table}[ht]
\centering
\caption{Outline of the major features of the five different Gmacs scenarios.} 
\label{tab:model_runs}
\begin{tabular}{lccc}
  \hline
  Scenario & Selectivity estimated & Use Francis LF weighting & Estimate $M_{1998}$ \\
  \hline
  Gmacs match   & No  & No  & Yes \\ 
  Gmacs base    & Yes & No  & Yes \\ 
  Gmacs M       & Yes & No  & No  \\ 
  Gmacs Francis & Yes & Yes & No  \\ 
  Gmacs force   & Yes & Yes & No  \\ 
  \hline
\end{tabular}
\end{table}


## Results

Results for all Gmacs scenarios are provided with comparisons to the 2015 model. We recommend the **Gmacs base** scenario for management purposes since it provides the best fit to the data and is most consistent with previous model specifications.

### a. Effective sample sizes and weighting factors.

Observed and estimated effective sample sizes are compared in Table \ref{tab:effn}. Effective sample sizes are also shown on size-composition plots (Figures \ref{fig:sc_pot}, \ref{fig:sc_pot_discarded}, and \ref{fig:sc_trawl_discarded}). 

Data weighting factors, SDNRs, and MARs are presented in Table \ref{tab:data_weighting}. The SDNR for the trawl survey is acceptable at 1.44 in the **Gmacs match** scenario, and improves to 1.41 in the **Gmacs base** scenario. In the **Gmacs M** model the SDNR of the trawl survey is slightly worse at 1.59, and is much worse in the exploratory **Gmacs force** scenario at 2.16. The SDNRs for the pot surveys show much the same pattern between each of the scenarios, but are much higher values (ranging from 3.95 to 5.19). These values are very high, and whilst they can be improved by down-weighting the pot survey, it is recommended that they be left as they are as the pot survey is one of the most important data series in this model. The MAR for the trawl and pot surveys shows the same pattern among each of the scenarios as the SDNR. The SDNR (and MAR) values for the trawl survey and pot survey size compositions were excellent, ranging from 0.78 to 1.30 (except for in the **Gmacs force** scenario where the weights were a little high). The SDNRs for the directed pot fishery size compositions are a little low, ranging from 0.64 to 0.79. However, the SDNRs (and MARs) were not used when weighting the size composition data sets in those scenarios that used the Francis weighting method (i.e. in the **Gmacs Francis**, and **Gmacs Force** scenarios). Instead, the Francis size composition weights were used (Francis 2011).

### b. Tables of estimates.

Model parameter estimates for each of the Gmacs scenarios are summarized in Tables \ref{tab:est_pars_match}, \ref{tab:est_pars_base}, \ref{tab:est_pars_Francis}, \ref{tab:est_pars_M}, and \ref{tab:est_pars_force}. These parameter estimates are compared in Table \ref{tab:est_pars_all}. Negative log-likelihood values and management measures for each of the Gmacs scenarios are compared in Tables \ref{tab:likelihood_components} and \ref{tab:management_quants}.

There is little difference in the parameter estimates within the **Gmacs match** and **Gmacs base** scenarios. This is reflected in the log-likelihood components and the management quantities. The parameter estimates in the **Gmacs M** scenario are a little different to the previous scenarios, particularly the estimate of the ADF&G pot survey catchability ($q$) (see Table \ref{tab:est_pars_all}).

### c. Graphs of estimates.

Estimated (and fixed) selectivities are compared in Figure \ref{fig:selectivity}.

The various model fits to total male ($>$ 89 mm CL) trawl survey biomass are compared in Figures \ref{fig:trawl_survey_biomass} and \ref{fig:trawl_survey_biomass_weights}. The fits to pot survey CPUE are compared in Figures \ref{fig:pot_survey_cpue} and \ref{fig:pot_survey_cpue_weights}. Standardized residuals of total male trawl survey biomass and pot survey CPUE are plotted in Figures \ref{fig:bts_resid_nmfs} and \ref{fig:bts_resid_adfg}. 

Fits to stage compositions for trawl survey, pot survey, and commercial observer data are shown in Figures \ref{fig:sc_pot}, \ref{fig:sc_pot_discarded}, and \ref{fig:sc_trawl_discarded} for the all scenarios. Bubble plots of stage composition residuals for trawl survey, pot survey, and commercial observer data are shown for the **Gmacs base**, **Gmacs M**, **Gmacs Francis**, and **Gmacs force** scenarios in Figures \ref{fig:sc_res_base}, \ref{fig:sc_res_Francis}, \ref{fig:sc_res_M}, and \ref{fig:sc_res_force}, respectively.

Fits to retained catch numbers and bycatch biomass are shown for all Gmacs scenarios in Figure \ref{fig:fit_to_catch}.

Estimated recruitment is compared in Figure \ref{fig:recruitment}. Estimated abundances by stage and mature male biomasses for all scenarios (including the 2015 model) are shown in Figures \ref{fig:init_N} and \ref{fig:mmb}. Estimated natural mortality each year ($M_t$) is presented in Figure \ref{fig:M_t}.

### d. Graphic evaluation of the fit to the data.

There is little difference between model estimated survey biomass in the gmacs scenarios when compared with the 2015 model (Figures \ref{fig:trawl_survey_biomass} and \ref{fig:pot_survey_cpue}). Looking at the model fits to the NMFS trawl survey biomass (Figure \ref{fig:trawl_survey_biomass}), the **Gmacs match** scenario is the most similar to the 2015 model, and the **Gmacs base** model is very similar as well. In all scenarios, Gmacs produces a better fit during the mid-late 1980s. However, since about 2010 Gmacs estimates a slighly lower survey biomass than the 2015 model in an attempt to better fit the ADF&G pot survey CPUE (Figure \ref{fig:pot_survey_cpue}). The three Gmacs scenarios that do not attempt to estimate natural mortality in 1998/99 (**Gmacs M**, **Gmacs Francis**, and **Gmacs force**) predict lower survey biomass from 1992 to 1998 than the other scenarios and the 2015 model. These same two runs also predict a lower survey biomass in recent years (since about 2010). While these two models may result in slightly worse fits to the data, they do not risk over-fitting the data in the same way the other scenarios do. As exptected the model that upweights the NMFS survey biomass and ADF&G pot survey CPUE (**Gmacs force**) provides a better fit to the survey biomass during the mid-late 1980s and a much better fit to the pot survey CPUE in the most recent two years (Figures \ref{fig:trawl_survey_biomass}, \ref{fig:trawl_survey_biomass_weights}, \ref{fig:pot_survey_cpue}, and \ref{fig:pot_survey_cpue_weights}). Keep in mind that this scenario was only included for exploratory purposes and forcing these weights resulted in worse SDNR and MAR values for the two abundance indices.

Estimated recruitment to the model is variable over time (Figure \ref{fig:recruitment}). Estimated recruitment during recent years is generally low in all scenarios. Estimated mature male biomass on 15 February also fluctuates strongly over time (Figure \ref{fig:mmb}).

### e. Retrospective and historic analyses.

Gmacs retrospective analyses under development.

### f. Uncertainty and sensitivity analyses.

Estimated standard deviations of parameters and selected management measures for the five Gmacs scenarios are summarized in Tables \ref{tab:est_pars_match}, \ref{tab:est_pars_base}, \ref{tab:est_pars_Francis}, \ref{tab:est_pars_M}, and \ref{tab:est_pars_force}. Probabilities for mature male biomass and OFL in 2016 are illustrated in Section F.

### g. Comparison of alternative model scenarios.

Both the **Gmacs match** and **Gmacs base** scenarios provide adequate matches between the 2015 model and its Gmacs equivalent. In fact, despite a few minor differences, estimates produced by the 2015 model are generally encompassed the in the uncertainty bounds of the **Gmacs match** model.

Looking at the plot of mature male biomass (Figure \ref{fig:mmb}), the **Gmacs force** scenario stands out as being quite different to the other models (including the 2015 model). This scenario results in a lower MMB from the mid-1908s through to the late-1990s, and is again lower in the most recent 5 years. This scenario upweights both the trawl survey and the pot survey abundance indices (it upweights the pot survey more than the trawl survey) and represents a model run that places greater trust in the abundance indices, particularly the pot survey, than other data sources.

Although the **Gmacs M** scenario presents a worse fit to the data, particularly the NMFS trawl-survey time series, this model does not simply allow a better fit to by estimating an unconstrained pulse in natural mortality. Although doing so produces a better fit to the model, it reduces predictive power and support for such a phenomena, anecdotal or otherwise, seems to be limited. It also raises concerns about what the implications would be for an “average” true natural mortality which can affect the management measures. Despite these concerns, more work is needed in the future to explore more parsimonious alternatives that provide better fits to the data.

In summary, we recommend the **Gmacs base** scenario for management purposes since it provides the best fit to the data and is most consistent with previous model specifications. Our initial preference was for **Gmacs M** since we had difficulty justifying an abrubt, single-year anomaly in natural mortality.  However, the fact that the residual pattern is worse and until further work can be completed on alternative model specifications (e.g., better accounting of spatial processes affecting the data), the **Gmacs base** model was considered reasonable and should be used for overfishing determination for this stock in 2016.


# F. Calculation of the OFL and ABC

The overfishing level (OFL) is the fishery-related mortality biomass associated with fishing mortality $F_\mathit{OFL}$. The SMBKC stock is currently managed as Tier 4 (2013 SAFE), and only a Tier 4 analysis is presented here. Thus given stock estimates or suitable proxy values of $B_\mathit{MSY}$ and $F_\mathit{MSY}$, along with two additional parameters $\alpha$ and $\beta$, $F_\mathit{OFL}$ is determined by the control rule
\begin{align}
    F_\mathit{OFL} &= 
    \begin{cases}
        F_\mathit{MSY}, &\text{ when } B/B_\mathit{MSY} > 1\\
        F_\mathit{MSY} \frac{\left( B/B_\mathit{MSY} - \alpha \right)}{(1 - \alpha)}, &\text{ when } \beta < B/B_\mathit{MSY} \le 1
    \end{cases}\\
    F_\mathit{OFL} &< F_\mathit{MSY} \text{ with directed fishery } F = 0 \text{ when } B/B_\mathit{MSY} \le \beta \notag
\end{align}
where $B$ is quantified as mature-male biomass (MMB) at mating with time of mating assigned a nominal date of 15 February. Note that as $B$ itself is a function of the fishing mortality $F_\mathit{OFL}$ (therefore numerical approximation of $F_\mathit{OFL}$ is required). As implemented for this assessment, all calculations proceed according to the model equations given in Appendix A. $F_\mathit{OFL}$ is taken to be full-selection fishing mortality in the directed pot fishery and groundfish trawl and fixed-gear fishing mortalities set at their model geometric mean values over years for which there are data-based estimates of bycatch-mortality biomass.

The currently recommended Tier 4 convention is to use the full assessment period, currently `r M[[2]]$spr_syr`-`r M[[2]]$spr_nyr`, to define a $B_\mathit{MSY}$ proxy in terms of average estimated MMB and to set $\gamma$ = 1.0 with assumed stock natural mortality $M$ = 0.18 $\text{yr}^{-1}$ in setting the $F_\mathit{MSY}$ proxy value $\gamma M$. The parameters $\alpha$ and $\beta$ are assigned their default values $\alpha$ = 0.10 and $\beta$ = 0.25. The $F_\mathit{OFL}$, OFL, ABC, and MMB in 2016 for all scenarios are summarized in Table \ref{tab:management_quants}. ABC is 80% of the OFL.
```{r management_quants, results = "asis"}
# df <- NULL
# for (ii in 2:4)
# {
#     x <- M[[ii]]
#     mmb <- x$ssb[length(x$ssb)]; names(mmb) <- paste0("$\\text{MMB}_{", x$mod_yrs[length(x$mod_yrs)], "}$")
#     fofl <- x$spr_fofl; names(fofl) <- "$F_\\text{OFL}$"
#     OFL <- x$spr_cofl; names(OFL) <- paste0("$\\text{OFL}_{", x$mod_yrs[length(x$mod_yrs)], "}$")
#     Bmsy <- x$spr_bspr; names(Bmsy) <- "$B_\\text{MSY}$"
#     ABC <- OFL * 0.8; names(ABC) <- paste0("$\\text{ABC}_{", x$mod_yrs[length(x$mod_yrs)], "}$")
#     v <- c(mmb, Bmsy, fofl, OFL, ABC)
#     df <- cbind(df, v)
# }
# df <- data.frame(rownames(df), df, row.names = NULL)
# names(df) <- c("Component","Gmacs base","Const M","Time-varying M")
# tab <- xtable(df, caption = "Comparisons of management measures for the five Gmacs model scenarios. Biomass and OFL are in tons.", label = "tab:management_quants", digits = 3)
# print(tab, caption.placement = "top", include.rownames = FALSE, sanitize.text.function = function(x){x})
```


# G. Rebuilding Analysis

This stock is not currently subject to a rebuilding plan.

# H. Data Gaps and Research Priorities

  1. Growth increments and molting probabilities as a function of size.
  2. Trawl survey catchability and selectivities.
  3. Temporal changes in spatial distributions near the island.
  4. Natural mortality.

# I. Projections and Future Outlook

With the decline of estimated population biomass during recent years, outlook for this stock is not promising. If the decline continues, the stock will fall to depleted status soon.

# J. Acknowledgements

We thank the Crab Plan Team, Doug Pengilly for reviewing the earlier draft of this manuscript. Some materials in the report are from the SAFE report prepared by Bill Gaeuman in 2014. We thank Andre Punt for his input into the Gmacs model and for finding the error in the old SMBKC model code.

# K. References

Alaska Department of Fish and Game (ADF&G). 2013. Crab observer training and deployment manual. Alaska Department of Fish and Game Shellfish Observer Program, Dutch Harbor. Unpublished.

Collie, J.S., A.K. Delong, and G.H. Kruse. 2005. Three-stage catch-survey analysis applied to blue king crabs. Pages 683-714 [In] Fisheries assessment and management in data-limited situations. University of Alaska Fairbanks, Alaska Sea Grant Report 05-02, Fairbanks.

Daly, B., R. Foy, and C. Armistead. 2014. The 2013 eastern Bering Sea continental shelf bottom trawl survey: results for commercial crab species. NOAA Technical Memorandum, NMFS-AFSC.

Donaldson, W.E., and S.C. Byersdorfer. 2005. Biological field techniques for lithodid crabs. University of Alaska Fairbanks, Alaska Sea Grant Report 05-03, Fairbanks.

Fitch, H., M. Deiman, J. Shaishnikoff, and K. Herring. 2012. Annual management report for the commercial and subsistence shellfish fisheries of the Bering Sea, 2010/11. Pages 75-1776 [In] Fitch, H., M. Schwenzfeier, B. Baechler, T. Hartill, M. Salmon, M. Deiman, E.

Evans, E. Henry, L. Wald, J. Shaishnikoff, K. Herring, and J. Wilson. 2012. Annual management report for the commercial and subsistence shellfish fisheries of the Aleutian Islands, Bering Sea and the Westward Region’s Shellfish Observer Program, 2010/11. Alaska Department of Fish and Game, Fishery Management Report No. 12-22, Anchorage.

Fournier, D.A., H.J. Skaug, J. Ancheta, J. Ianelli, A. Magnusson, M.N. Maunder, A. Nielsen, and J. Sibert. 2012. AD Model Builder: using automatic differentiation for statistical inference of highly parameterized complex nonlinear models. Optim. Methods Softw. 27:233-249.

Francis, R.I.C.C. 2011. Data weighting in statistical fisheries stock assessment models. Can. J. Fish. Aquat. Sci. 68: 1124-1138.

Gaeuman, W.B. 2013. Summary of the 2012/13 mandatory crab observer program database for the Bering Sea/Aleutian Islands commercial crab fisheries. Alaska Department of Fish and Game, Fishery Data Series No. 13-54, Anchorage. 

Gish, R.K., V.A. Vanek, and D. Pengilly. 2012. Results of the 2010 triennial St. Matthew Island blue king crab pot survey and 2010/11 tagging study. Alaska Department of Fish and Game, Fishery Management Report No. 12-24, Anchorage.

Jensen, G.C. and D.A. Armstrong. 1989. Biennial reproductive cycle of blue king crab, Paralithodes platypus, at the Pribilof Islands, Alaska and comparison to a congener, P. camtschatica. Can. J. Fish. Aquat. Sci. 46: 932-940.

Moore, H., L.C. Byrne, and D. Connolly. 2000. Alaska Department of Fish and Game summary of the 1998 mandatory shellfish observer program database. Alaska Dept. Fish and Game, Commercial Fisheries Division, Reg. Inf. Rep. 4J00-21, Kodiak.

North Pacific Fishery Management Council (NPFMC). 1998. Fishery Management Plan for Bering Sea/Aleutian Islands king and Tanner crabs. North Pacific Fishery Management Council, Anchorage.

North Pacific Fishery Management Council (NPFMC). 1999. Environmental assessment/regulatory impact review/initial regulatory flexibility analysis for Amendment 11 to the Fishery Management Plan for Bering Sea/Aleutian Islands king and Tanner crabs. North Pacific Fishery Management Council, Anchorage.

North Pacific Fishery Management Council (NPFMC). 2000. Environmental assessment/regulatory impact review/initial regulatory flexibility analysis for proposed Amendment 15 to the Fishery Management Plan for king and Tanner crab fisheries in the Bering Sea/Aleutian Islands and regulatory amendment to the Fishery Management Plan for the groundfish fishery of the Bering Sea and Aleutian Islands area: A rebuilding plan for the St. Matthew blue king crab stock. North Pacific Fishery Management Council, Anchorage. Draft report.

North Pacific Fishery Management Council (NPFMC). 2007. Public Review Draft: Environmental assessment for proposed Amendment 24 to the Fishery Management Plan for Bering Sea and Aleutian Islands king and Tanner crabs to revise overfishing definitions. 14 November 2007. North Pacific Fishery Management Council, Anchorage.

Otto, R.S. 1990. An overview of eastern Bering Sea king and Tanner crab fisheries. Pages 9-26 [In] Proceedings of the international symposium on king and Tanner crabs. University of Alaska Fairbanks, Alaska Sea Grant Program Report 90-4, Fairbanks.

Otto, R.S., and P.A. Cummiskey. 1990. Growth of adult male blue king crab (Paralithodes platypus). Pages 245-258 [In] Proceedings of the international symposium on king and Tanner crabs. University of Alaska Fairbanks, Alaska Sea Grant Report 90-4, Fairbanks.

Paul, J.M., A. J. Paul, R.S. Otto, and R.A. MacIntosh. 1991. Spermatophore presence in relation to carapace length for eastern Bering Sea blue king crab (Paralithodes platypus, Brandt, 1850) and red king crab (P. Camtschaticus, Tilesius, 1815). J. Shellfish Res. 10: 157-163.

Pengilly, D. and D. Schmidt. 1995. Harvest Strategy for Kodiak and Bristol Bay Red king Crab and St. Matthew Island and Pribilof Blue King Crab. Alaska Department of Fish and Game, Commercial Fisheries Management and Development Division, Special Publication Number 7, Juneau.

Schirripa, M.J., C.P. Goodyear, and R.M. Methot. 2009. Testing different methods of incorporating climate data into the assessment of US West Coast sablefish. ICES Journal of Marine Science, 66: 1605–1613. 

Somerton, D.A., and R.A. MacIntosh. 1983. The size at sexual maturity of blue king crab, Paralithodes platypus, in Alaska. Fishery Bulletin 81: 621-828.

Wilderbuer, T., D. G. Nichol, and J. Ianelli. 2013. Assessment of the yellowfin sole stock in the Bering Sea and Aleutian Islands. Pages 619-708 in 2013 North Pacific Groundfish Stock Assessment and Fishery Evaluation Reports for 2014. North Pacific Fishery Management Council, Anchorage.

Zheng, J. 2005. A review of natural mortality estimation for crab stocks: data-limited for every stock? Pages 595-612 [In] Fisheries Assessment and Management in Data-Limited Situations. University of Alaska Fairbanks, Alaska Sea Grant Program Report 05-02, Fairbanks.

Zheng, J., and G.H. Kruse. 2002. Assessment and management of crab stocks under uncertainty of massive die-offs and rapid changes in survey catchability. Pages 367-384 [In] A.J. Paul,E.G. Dawe, R. Elner, G.S. Jamieson, G.H. Kruse, R.S. Otto, B. Sainte-Marie, T.C. Shirley, and D. Woodby (eds.). Crabs in Cold Water Regions: Biology, Management, and Economics. University of Alaska Fairbanks, Alaska Sea Grant Report 02-01, Fairbanks.

Zheng, J., M.C. Murphy, and G.H. Kruse. 1997. Application of catch-survey analysis to blue king crab stocks near Pribilof and St. Matthew Islands. Alaska Fish. Res. Bull. 4:62-74.


\newpage\clearpage

```{r length_weight, results = "asis"}
df <- data.frame(M[[2]]$mod_yrs, M[[2]]$mean_wt[,3:18] * 1000)
#xxxx
names(df) <- c("Year","Stage-1","Stage-2","Stage-3")
#tab <- xtable(df, caption = "Mean weight (kg) by stage in used in all of the models (provided as a vector of weights at length each year to Gmacs).", label = "tab:length_weight", digits = c(0,0,1,1,1))
#print(tab, caption.placement = "top", include.rownames = FALSE, sanitize.text.function = function(x){x})
```

\begin{table}[ht]
\centering
\caption{Observed and assumed sample sizes for observer data from the directed pot fishery, the NMFS trawl survey, and the ADF\&G pot survey.} 
\label{tab:effn}
\begin{tabular}{lccccccc}
  \hline
  & \multicolumn{3}{c}{Observed sample sizes} & \multicolumn{3}{c}{Assumed sample sizes} \\
  \cline{2-4}\cline{6-8}
  Year & Observer pot & NMFS trawl & ADF\&G pot & & Observer pot & NMFS trawl & ADF\&G pot \\ 
  \hline
  1978 &              & 157        &          & &              & 50         & \\
  1979 &              & 178        &          & &              & 50         & \\
  1980 &              & 185        &          & &              & 50         & \\
  1981 &              & 140        &          & &              & 50         & \\
  1982 &              & 271        &          & &              & 50         & \\
  1983 &              & 231        &          & &              & 50         & \\
  1984 &              & 105        &          & &              & 50         & \\
  1985 &              &  93        &          & &              & 46.5       & \\
  1986 &              &  46        &          & &              & 23         & \\
  1987 &              &  71        &          & &              & 35.5       & \\
  1988 &              &  81        &          & &              & 40.5       & \\
  1989 &              & 208        &          & &              & 50         & \\
  1990 &  150         & 170        &          & & 15           & 50         & \\
  1991 &  3393        & 197        &          & & 25           & 50         & \\
  1992 &  1606        & 220        &          & & 25           & 50         & \\
  1993 &  2241        & 324        &          & & 25           & 50         & \\
  1994 &  4735        & 211        &          & & 25           & 50         & \\
  1995 &  663         & 178        &  4624    & & 25           & 50         & 100 \\
  1996 &  489         & 285        &          & & 25           & 50         & \\
  1997 &  3195        & 296        &          & & 25           & 50         & \\
  1998 &  1323        & 243        &  4812    & & 25           & 50         & 100 \\
  1999 &              &  52        &          & &              & 26         & \\
  2000 &              &  61        &          & &              & 30.5       & \\
  2001 &              &  91        &  3255    & &              & 45.5       & 100 \\
  2002 &              &  38        &          & &              & 19         & \\
  2003 &              &  65        &          & &              & 32.5       & \\
  2004 &              &  48        &  640     & &              & 24         & 100 \\
  2005 &              &  42        &          & &              & 21         & \\
  2006 &              & 126        &          & &              & 50         & \\
  2007 &              & 250        &  3319    & &              & 50         & 100 \\
  2008 &              & 167        &          & &              & 50         & \\
  2009 &  19802       & 251        &          & & 50           & 50         & \\
  2010 &  45466       & 388        &  3920    & & 50           & 50         & 100 \\
  2011 &  58667       & 318        &          & & 50           & 50         & \\
  2012 &  57282       & 193        &          & & 50           & 50         & \\
  2013 &              &  74        &  2167    & &              & 37         & 100 \\
  2014 &  9906        & 181        &          & & 50           & 50         & \\
  2015 &  3248        & 153        &  1077    & & 50           & 50         & 100 \\
  2016 &              & 108        &   777    & &              & 50         & 100 \\
  \hline
\end{tabular}
\end{table}

```{r est_pars_2, results = "asis"}
# x <- M[[2]]$fit
# i <- c(grep("m_dev", x$names)[1],
#        grep("theta", x$names),
#        grep("survey_q", x$names),
#        grep("log_fbar", x$names),
#        grep("sd_fofl", x$names),
#        grep("sd_ofl", x$names)
#        )
# Parameter <- x$names[i]
# Estimate <- x$est[i]
# SD <- x$std[i]
# j <- grep("survey_q", Parameter)
# Estimate[j] <- Estimate[j] * 1000
# SD[j] <- SD[j] * 1000
# #Parameter <- c("Natural mortality deviation in 1998/99 ($\\delta^M_{1998})$","$\\log (\\bar{R})$","$\\log (n^0_1)$","$\\log (n^0_2)$","$\\log (n^0_3)$","$\\log (\\bar{F}^\\text{df})$","$\\log (\\bar{F}^\\text{tb})$","$\\log (\\bar{F}^\\text{fb})$","$F_\\text{OFL}$","OFL")
# df <- data.frame(Parameter, Estimate, SD)
# tab <- xtable(df, caption = "Model parameter estimates, selected derived quantities, and their standard deviations (SD) for the {\\bf Gmacs match} model.", label = "tab:est_pars_match", digits = 3)
# print(tab, caption.placement = "top", include.rownames = FALSE, sanitize.text.function = function(x){x})
```


\newpage\clearpage

```{r molt_prob, fig.cap = "Comparisons of the estimated (and fixed to match the 2015 model selectivities in the Gmacs base scenario) stage-1 and stage-2 selectivities for each of the different model scenarios (the stage-3 selectivities are all fixed at 1). Estimated selectivities are shown for the directed pot fishery, the trawl bycatch fishery, the fixed bycatch fishery, the NMFS trawl survey, and the ADF&G pot survey. Two selectivity periods are estimated in the directed pot fishery, from 1978-2008 and 2009-2016.\\label{fig:molt_prob}"}
mp1 <- c(0.999966, 0.999934, 0.99987, 0.999746, 0.999504, 0.999031, 0.998109, 0.996311, 0.992816, 0.986056, 0.973108, 0.948763, 0.90454, 0.829026, 0.712748, 0.559415, 0.393843, 0.249522, 0.145401, 0.0800909)
mp2 <- c(0.999292, 0.998841, 0.998102, 0.996894, 0.994921, 0.991705, 0.986481, 0.978039, 0.964516, 0.943149, 0.910114, 0.86072, 0.790431, 0.697152, 0.584195, 0.461642, 0.343556, 0.242093, 0.163148, 0.106334)

M[[2]]$molt_probability[1:5,] <- matrix(mp1, nrow = 5, ncol = 20, byrow = TRUE)
M[[2]]$molt_probability[6:42,] <- matrix(mp2, nrow = 37, ncol = 20, byrow = TRUE)
M[[2]]$molt_probability[43:84,] <- 1

plot_molt_prob(M)
```

```{r growth_inc, fig.cap = "Comparisons of the estimated (and fixed to match the 2015 model selectivities in the Gmacs base scenario) stage-1 and stage-2 selectivities for each of the different model scenarios (the stage-3 selectivities are all fixed at 1). Estimated selectivities are shown for the directed pot fishery, the trawl bycatch fishery, the fixed bycatch fishery, the NMFS trawl survey, and the ADF&G pot survey. Two selectivity periods are estimated in the directed pot fishery, from 1978-2008 and 2009-2016.\\label{fig:growth_inc}"}
gi1 <- c(13.8, 12.2, 10.5, 8.4, 7.5, 7, 6.6, 6.1, 5.6, 5.1, 4.6, 4.1, 3.6, 3.2, 2.7, 2.2, 1.7, 1.2, 0.7, 0.4)
gi2 <- c(15.4, 13.8, 12.2, 10.5, 8.9, 7.9, 7.2, 6.6, 6.1, 5.6, 5.1, 4.6, 4.1, 3.6, 3.2, 2.7, 2.2, 1.7, 1.2, 0.7)
gi3 <- c(15.1, 14, 12.9, 11.8, 10.6, 8.7, 7.4, 6.6, 6.1, 5.6, 5.1, 4.6, 4.1, 3.6, 3.2, 2.7, 2.2, 1.7, 1.2, 0.7)
gi4 <- c(16.5, 16.5, 16.4, 16.3, 16.3, 16.2, 16.2, 16.1, 16.1, 16, 16, 15.9, 15.8, 15.8, 15.7, 15.7, 15.6, 15.6, 15.5, 15.5)

M1 <- M
M1[[3]] <- M[[2]]
M1[[4]] <- M[[2]]
names(M1) <- c('Gmacs','1975-1982','1983-1993','1994-present')

M1[[2]]$pMoltInc[M1[[2]]$iMoltIncSex == 1] <- gi4
M1[[2]]$pMoltInc[M1[[2]]$iMoltIncSex == 2] <- gi1
M1[[3]]$pMoltInc[M1[[3]]$iMoltIncSex == 1] <- gi4
M1[[3]]$pMoltInc[M1[[3]]$iMoltIncSex == 2] <- gi2
M1[[4]]$pMoltInc[M1[[4]]$iMoltIncSex == 1] <- gi4
M1[[4]]$pMoltInc[M1[[4]]$iMoltIncSex == 2] <- gi3

plot_growth_inc(M1)
```

```{r growth_trans, fig.cap = "Probability of growth transition by stage. Each of the panels represent the stage before a transition. The x-axes represent the stage after a transition. The size transition matrix was provided as an input directly to Gmacs (as it was during the 2015 SMBKC assessment).\\label{fig:growth_trans}", fig.height = 15}
gm_f1 <- matrix(scan(text = "
 0 0.0711234 0.398022 0.372157 0.132835 0.0258631 0 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0.145003 0.475426 0.294412 0.0762733 0.0088857 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0.2719 0.496055 0.195229 0.0357593 0.00105686 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0.0152297 0.486895 0.400566 0.087935 0.0093735 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0.0276419 0.580508 0.332256 0.0563255 0.00326825 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0.0648859 0.613229 0.2845 0.0373848 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0.0847572 0.640095 0.249846 0.0253018 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0.115808 0.663216 0.207271 0.0137048 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0.171288 0.66037 0.164023 0.00431962 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0.228945 0.647251 0.123804 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0.298984 0.615451 0.0855653 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0.380675 0.565457 0.0538671 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0.471491 0.499965 0.0285444 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0.549292 0.440929 0.00977904 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0.645044 0.354956 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0
"), nrow = 20, ncol = 20, byrow = TRUE)

gm_f2 <- matrix(scan(text = "
 0 0.0315365 0.29835 0.415939 0.203388 0.0507866 0 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0.0715876 0.40062 0.374586 0.133701 0.0195053 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0.1459 0.478366 0.296233 0.076745 0.0027561 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0.00293279 0.2747 0.495812 0.195133 0.0314218 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0.0106724 0.435786 0.435296 0.110078 0.00816757 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0.0342176 0.540301 0.364189 0.0612922 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0.0569465 0.602618 0.304312 0.0361236 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0.0853184 0.644334 0.2515 0.0188471 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0.12638 0.660964 0.206567 0.00608825 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0.172781 0.666124 0.161095 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0.231245 0.653752 0.115003 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0.302451 0.622588 0.0749609 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0.385785 0.573046 0.041169 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0.478849 0.507767 0.0133836 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0.560249 0.439751 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0
"), nrow = 20, ncol = 20, byrow = TRUE)

gm_m1 <- matrix(scan(text = "
 0 0.0370031 0.317399 0.410924 0.1895 0.0451734 0 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0.0650656 0.388977 0.382571 0.142076 0.0213101 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0.108856 0.451675 0.335511 0.100072 0.00388567 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0.000901521 0.173397 0.493689 0.275204 0.0568086 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0.00272488 0.269398 0.503933 0.203784 0.0201606 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0.0184362 0.463497 0.428655 0.0894116 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0.0495539 0.587504 0.322594 0.040348 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0.0853184 0.644334 0.2515 0.0188471 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0.12638 0.660964 0.206567 0.00608825 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0.172781 0.666124 0.161095 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0.231245 0.653752 0.115003 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0.302451 0.622588 0.0749609 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0.385785 0.573046 0.041169 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0.478849 0.507767 0.0133836 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0.560249 0.439751 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0
"), nrow = 20, ncol = 20, byrow = TRUE)

gm_m2 <- matrix(scan(text = "
 0 0.0374249 0.257557 0.369321 0.236624 0.0941279 0.00494521 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0.0374249 0.257557 0.369321 0.236624 0.0941279 0.00494521 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0.0390317 0.262221 0.369126 0.233146 0.0916878 0.00478779 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0.0406974 0.266902 0.368829 0.229653 0.0892849 0.00463398 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0.0406974 0.266902 0.368829 0.229653 0.0892849 0.00463398 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0.0424237 0.271597 0.36843 0.226146 0.0869192 0.00448375 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0.0424237 0.271597 0.36843 0.226146 0.0869192 0.00448375 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0.0442121 0.276304 0.367927 0.222629 0.084591 0.00433705 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0.0442121 0.276304 0.367927 0.222629 0.084591 0.00433705 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0.0460642 0.28102 0.367321 0.219101 0.0823004 0.00419385 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0.0460642 0.28102 0.367321 0.219101 0.0823004 0.00419385 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0.0479817 0.285742 0.36661 0.215565 0.0800475 0.00405411 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0.0499661 0.290466 0.365796 0.212022 0.0778325 0.0039178 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0.0499661 0.290466 0.365796 0.212022 0.0778325 0.0039178
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0.0520191 0.295189 0.364877 0.208474 0.0794404
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0.0520191 0.295189 0.364877 0.287915
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0.0541422 0.29991 0.645948
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0.0541422 0.945858
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1
"), nrow = 20, ncol = 20, byrow = TRUE)

M1 <- M
M1[[3]] <- M[[2]]
names(M1) <- c('Gmacs','1983-93','1994-present')

M1[[2]]$growth_transition[1:20,] <- gm_m1
M1[[2]]$growth_transition[21:40,] <- gm_f1
M1[[3]]$growth_transition[1:20,] <- gm_m2
M1[[3]]$growth_transition[21:40,] <- gm_f2

plot_growth_transition(M1, xlab = "Carapace width after transition (mm)")
```

```{r size_trans, fig.cap = "Probability of size transition by stage (i.e. the combination of the growth matrix and molting probabilities). Each of the panels represent the stage before a transition. The x-axes represent the stage after a transition. The size transition matrix was provided as an input directly to Gmacs (as it was during the 2015 SMBKC assessment).\\label{fig:size_trans}", fig.height = 15}
gm_f1 <- matrix(scan(text = "
 0 0.0711234 0.398022 0.372157 0.132835 0.0258631 0 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0.145003 0.475426 0.294412 0.0762733 0.0088857 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0.2719 0.496055 0.195229 0.0357593 0.00105686 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0.0152297 0.486895 0.400566 0.087935 0.0093735 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0.0276419 0.580508 0.332256 0.0563255 0.00326825 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0.0648859 0.613229 0.2845 0.0373848 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0.0847572 0.640095 0.249846 0.0253018 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0.115808 0.663216 0.207271 0.0137048 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0.171288 0.66037 0.164023 0.00431962 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0.228945 0.647251 0.123804 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0.298984 0.615451 0.0855653 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0.380675 0.565457 0.0538671 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0.471491 0.499965 0.0285444 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0.549292 0.440929 0.00977904 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0.645044 0.354956 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0
"), nrow = 20, ncol = 20, byrow = TRUE)

gm_f2 <- matrix(scan(text = "
 0 0.0315365 0.29835 0.415939 0.203388 0.0507866 0 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0.0715876 0.40062 0.374586 0.133701 0.0195053 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0.1459 0.478366 0.296233 0.076745 0.0027561 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0.00293279 0.2747 0.495812 0.195133 0.0314218 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0.0106724 0.435786 0.435296 0.110078 0.00816757 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0.0342176 0.540301 0.364189 0.0612922 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0.0569465 0.602618 0.304312 0.0361236 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0.0853184 0.644334 0.2515 0.0188471 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0.12638 0.660964 0.206567 0.00608825 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0.172781 0.666124 0.161095 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0.231245 0.653752 0.115003 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0.302451 0.622588 0.0749609 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0.385785 0.573046 0.041169 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0.478849 0.507767 0.0133836 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0.560249 0.439751 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0
"), nrow = 20, ncol = 20, byrow = TRUE)

gm_m1 <- matrix(scan(text = "
 0 0.0370031 0.317399 0.410924 0.1895 0.0451734 0 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0.0650656 0.388977 0.382571 0.142076 0.0213101 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0.108856 0.451675 0.335511 0.100072 0.00388567 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0.000901521 0.173397 0.493689 0.275204 0.0568086 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0.00272488 0.269398 0.503933 0.203784 0.0201606 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0.0184362 0.463497 0.428655 0.0894116 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0.0495539 0.587504 0.322594 0.040348 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0.0853184 0.644334 0.2515 0.0188471 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0.12638 0.660964 0.206567 0.00608825 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0.172781 0.666124 0.161095 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0.231245 0.653752 0.115003 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0.302451 0.622588 0.0749609 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0.385785 0.573046 0.041169 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0.478849 0.507767 0.0133836 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0.560249 0.439751 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0
"), nrow = 20, ncol = 20, byrow = TRUE)

gm_m2 <- matrix(scan(text = "
 0 0.0374249 0.257557 0.369321 0.236624 0.0941279 0.00494521 0 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0.0374249 0.257557 0.369321 0.236624 0.0941279 0.00494521 0 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0.0390317 0.262221 0.369126 0.233146 0.0916878 0.00478779 0 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0.0406974 0.266902 0.368829 0.229653 0.0892849 0.00463398 0 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0.0406974 0.266902 0.368829 0.229653 0.0892849 0.00463398 0 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0.0424237 0.271597 0.36843 0.226146 0.0869192 0.00448375 0 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0.0424237 0.271597 0.36843 0.226146 0.0869192 0.00448375 0 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0.0442121 0.276304 0.367927 0.222629 0.084591 0.00433705 0 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0.0442121 0.276304 0.367927 0.222629 0.084591 0.00433705 0 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0.0460642 0.28102 0.367321 0.219101 0.0823004 0.00419385 0 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0.0460642 0.28102 0.367321 0.219101 0.0823004 0.00419385 0 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0.0479817 0.285742 0.36661 0.215565 0.0800475 0.00405411 0 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0.0499661 0.290466 0.365796 0.212022 0.0778325 0.0039178 0
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0.0499661 0.290466 0.365796 0.212022 0.0778325 0.0039178
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0.0520191 0.295189 0.364877 0.208474 0.0794404
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0.0520191 0.295189 0.364877 0.287915
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0.0541422 0.29991 0.645948
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0.0541422 0.945858
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1
 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 0 1
"), nrow = 20, ncol = 20, byrow = TRUE)

M1 <- M
M1[[3]] <- M[[2]]
names(M1) <- c('Gmacs','1983-93','1994-present')

#M1[[2]]$growth_transition[1:20,] <- gm_m1
#M1[[2]]$growth_transition[21:40,] <- gm_f1
#M1[[3]]$growth_transition[1:20,] <- gm_m2
#M1[[3]]$growth_transition[21:40,] <- gm_f2

M1[[2]]$size_transition_M <- gm_m1
M1[[2]]$size_transition_F <- gm_f1
M1[[3]]$size_transition_M <- gm_m2
M1[[3]]$size_transition_F <- gm_f2

#plot_size_transition(M1, xlab = "Carapace width after transition (mm)")
```

```{r selectivity, fig.cap = "Comparisons of the estimated (and fixed to match the 2015 model selectivities in the Gmacs base scenario) stage-1 and stage-2 selectivities for each of the different model scenarios (the stage-3 selectivities are all fixed at 1). Estimated selectivities are shown for the directed pot fishery, the trawl bycatch fishery, the fixed bycatch fishery, the NMFS trawl survey, and the ADF&G pot survey. Two selectivity periods are estimated in the directed pot fishery, from 1978-2008 and 2009-2016.\\label{fig:selectivity}", fig.height = 15}
plot_selectivity(M, ncol = 4)
```

```{r trawl_survey_biomass, fig.cap = "Comparisons of area-swept estimates of total male survey biomass (tons) and model predictions for the 2015 model and each of the Gmacs model scenarios. The error bars are plus and minus 2 standard deviations.\\label{fig:trawl_survey_biomass}"}
A <- M
#A[[1]]$obs_cpue <- ? as per usual Jie outputs need to be customized
plot_cpue(A, subsetby = "NMFS Trawl", ylab = "Survey biomass (tons)")
```

```{r trawl_survey_biomass_weights, fig.cap = "Comparisons of area-swept estimates of total male survey biomass (tons) and model predictions for the 2015 model and each of the Gmacs model scenarios. The solid black error bars are plus and minus 2 standard deviations derived using the original survey CVs. The dotted error bars are plus and minus 2 standard deviations but represent the weighted survey CVs.\\label{fig:trawl_survey_biomass_weights}"}
plot_cpue(M, subsetby = "NMFS Trawl", ylab = "Survey biomass (tons)", ShowEstErr = TRUE)
```

\newpage\clearpage

```{r pot_survey_cpue, fig.cap = "Comparisons of total male pot survey CPUEs and model predictions for the 2015 model and each of the Gmacs model scenarios. The error bars are plus and minus 2 standard deviations.\\label{fig:pot_survey_cpue}"}
#plot_cpue(M, "BSFRF", ylab = "Pot survey CPUE (crab/potlift)")
```

```{r pot_survey_cpue_weights, fig.cap = "Comparisons of total male pot survey CPUEs and model predictions for the 2015 model and each of the Gmacs model scenarios. The solid black error bars are plus and minus 2 standard deviations derived using the original survey CVs. The dotted error bars are plus and minus 2 standard deviations but represent the weighted survey CVs.\\label{fig:pot_survey_cpue_weights}"}
#plot_cpue(M, "BSFRF", ylab = "Pot survey CPUE (crab/potlift)", ShowEstErr = TRUE)
```

```{r bts_resid_nmfs, fig.cap = "Standardized residuals for area-swept estimates of total male survey biomass for each of the Gmacs model scenarios. \\label{fig:bts_resid_nmfs}"}
A <- M; A[[jj]] <- NULL
plot_cpue_res(A, "NMFS Trawl")
```

\newpage\clearpage

```{r sc_pot, fig.cap = "Observed and model estimated size-frequencies of SMBKC by year retained in the directed pot fishery for the 2015 model and each of the Gmacs model scenarios. Note that there is no model estimated size-frequency for the 2015 model during the 2015 year.\\label{fig:sc_pot}"}
#plot_size_comps(M, 1)
```

```{r sc_pot_discarded, fig.cap = "Observed and model estimated size-frequencies of discarded male SMBKC by year in the NMFS trawl survey for the 2015 model and each of the Gmacs model scenarios. Note that there is no model estimated size-frequency for the 2015 model during the 2016 year.\\label{fig:sc_pot_discarded}"}
#plot_size_comps(M, 2)
```

```{r sc_trawl_discarded, fig.cap = "Observed and model estimated size-frequencies of discarded SMBKC by year in the ADF&G pot survey for the 2015 model and each of the Gmacs model scenarios. Note that there is no model estimated size-frequency for the 2015 model during the 2016 year.\\label{fig:sc_trawl_discarded}"}
#plot_size_comps(M, 3)
```


\newpage\clearpage

```{r fit_to_catch, fig.cap = "Comparison of observed and model predicted retained catch and bycatches in each of the Gmacs models. Note that difference in units between each of the panels, some panels are expressed in numbers of crab, some as biomass (tons).\\label{fig:fit_to_catch}", fig.height = 12}
A <- M; A[[jj]] <- NULL
plot_catch(A)
```

```{r recruitment, fig.cap = "Comparisons of estimated recruitment time series during 1979-2016 in each of the scenarios. The solid horizontal lines in the background represent the estimate of the average recruitment parameter ($\\bar{R}$) in each model scenario.\\label{fig:recruitment}", fig.height = 12}
# estimated number of recruitments female: seq(1968,2008)
# 2016
#rec_f <- c(17.0707, 25.925, 29.8683, 51.6996, 38.3764, 14.6497, 70.4083, 33.2085, 52.9009, 4.57629, 27.8906, 6.15353, 4.37952, 4.13648, 11.4324, 6.96502, 0.718979, 4.5763, 0.575209, 29.1253, 1.99369, 0.763189, 4.95964, 16.7755, 5.46132, 6.11056, 30.5387, 4.0503, 7.25958, 31.4988, 5.92376, 5.27457, 5.05108, 5.89551, 8.26562, 8.63609, 5.11815, 2.52513, 1.03167, 2.53766, 1.56217)
# 2017
rec_f <- c(16.6203, 24.8987, 28.6467, 49.4034, 36.5597, 14.0565, 68.6874, 32.5652, 52.1421, 4.56213, 27.6828, 6.11474, 4.35065, 4.10836, 11.3223, 6.87334, 0.736218, 4.50874, 0.594796, 28.8167, 2.00399, 0.774134, 4.90323, 16.59, 5.40813, 6.04792, 30.0304, 3.99239, 7.09898, 30.7368, 5.7649, 5.15361, 4.91422, 5.75793, 8.07117, 8.37412, 5.00154, 2.47013, 1.01736, 2.43727, 1.55705)
# estimated number of recruitments male: seq(1968,2008)
# 2016
#rec_m <- c(17.1233, 15.2388, 18.748, 24.6463, 29.9857, 13.3639, 64.4212, 32.6258, 33.2142, 4.03989, 15.6044, 6.4734, 3.24033, 3.73295, 11.7251, 7.40212, 1.2922, 5.87057, 0.914705, 28.1729, 4.53894, 1.90793, 6.73314, 15.415, 5.98756, 3.11696, 24.1102, 4.20828, 8.77771, 22.1832, 12.1232, 6.86644, 4.34433, 4.41878, 8.04963, 7.49265, 5.47651, 4.54582, 1.23435, 2.84266, 1.57952)
# 2017
rec_m <- c(16.8336, 14.7309, 18.1631, 23.9778, 29.2141, 12.9747, 63.1229, 32.1647, 33.0318, 3.99647, 15.5154, 6.42593, 3.20255, 3.68911, 11.5983, 7.32998, 1.27599, 5.80391, 0.899527, 27.912, 4.52506, 1.87739, 6.66915, 15.3099, 5.95461, 3.07542, 23.9416, 4.16347, 8.68231, 21.91, 11.9311, 6.74721, 4.24616, 4.32773, 7.8643, 7.28119, 5.34104, 4.44557, 1.19985, 2.7685, 1.52714)
A <- M
ii <- which(A[[2]]$fit$names %in% "sd_log_recruits")#; ii <- ii[length(ii)]
A[[2]]$fit$est[ii] <- log(c(rec_m, NA, rec_f, NA) * 1e6)
A[[2]]$fit$std[ii] <- NA
plot_recruitment(A)
```

```{r mature_male_biomass, fig.cap = "Comparisons of estimated mature male biomass (MMB) time series on 15 February during 1978-2016 for each of the model scenarios.\\label{fig:mmb}"}
ii <- M[[2]]$fit$names == "sd_log_ssb"
M[[2]]$fit$est[ii] <- log(c(80955.7, 90716.6, 94074.4, 97136, 82946.7, 24783.7, 8313.41, 8157.77, 8593.47, 6490.84, 10386, 15126.5, 21635.7, 27635.8, 31316.5, 29177.6, 24245.2, 22123.5, 19574, 25045, 27809.1, 25767.2, 23897.7, 26161.7, 30664, 30535.2, 29357.2, 31147.2, 29741.6, 27613.8, 27847.8, 29785.6, 27042, 28194.5, 31980.5, 32224.4, 32349.9, 31140, 30002.1, 28843, 27680.4, 23940))
M[[2]]$fit$std[ii] <- NA
plot_ssb(M, ylab = "Mature male biomass (tons) on 15 February")
```

```{r init_rec, fig.cap = "Distribution of carapace width (mm) at recruitment.\\label{fig:init_rec}"}
A <- M
A[[3]] <- A[[2]]
names(A) <- c("Gmacs","Females 2016","Males 2016")
# estimated proportions of recruitments by length
# 2016
#A[[2]]$rec_sdd <- c(0.252677, 0.347303, 0.25851, 0.111753, 0.029757, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
#A[[3]]$rec_sdd <- c(0.251349, 0.343023, 0.256401, 0.112449, 0.0306581, 0.00545854, 0.000662105, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
# 2017
A[[2]]$rec_sdd <- c(0.252692, 0.34716, 0.258476, 0.111844, 0.0298277, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
A[[3]]$rec_sdd <- c(0.251382, 0.342222, 0.256174, 0.112911, 0.0310369, 0.00558752, 0.000687101, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
plot_recruitment_size(A, xlab = "Carapace width (mm)")
```

```{r init_N, fig.cap = "Numbers by stage each year (at the beginning of the model year, i.e. 1 July, season 1) in each of the models including the 2015 model.\\label{fig:init_N}", fig.height = 15}
plot_numbers(M)
```

```{r natural_mortality, fig.cap = "Time-varying natural mortality ($M_t$). Estimated pulse period occurs in 1998/99 (i.e. $M_{1998}$). \\label{fig:M_t}"}
# Natural mortality for males:
# 2016
#Mm <- c(0.18, 0.270878, 0.270878, 0.270878, 0.270878, 0.986707, 0.986707, 0.986707, 0.986707, 0.986707, 0.270878, 0.270878, 0.270878, 0.270878, 0.270878, 0.270878, 0.270878, 0.270878, 0.270878, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18)
#Mf <- c(0.18, 0.18, 0.18, 0.18, 0.18, 0.64029, 0.64029, 0.64029, 0.64029, 0.64029, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18)
# 2017
Mm <- c(0.18, 0.271176, 0.271176, 0.271176, 0.271176, 0.979748, 0.979748, 0.979748, 0.979748, 0.979748, 0.271176, 0.271176, 0.271176, 0.271176, 0.271176, 0.271176, 0.271176, 0.271176, 0.271176, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18)
Mf <- c(0.18, 0.18, 0.18, 0.18, 0.18, 0.637793, 0.637793, 0.637793, 0.637793, 0.637793, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18, 0.18)
A <- M
A[[2]]$M[1:42,] <- Mm
A[[2]]$M[43:84,] <- Mf
plot_natural_mortality(A, knots = NULL, slab = "Model")
```


\newpage\clearpage

# Appendix A: SMBKC Model Description

## 1. Introduction

The Gmacs model has been specified to account only for male crab at least 90 mm in carapace length (CL). These are partitioned into three stages (size-classes) determined by CL measurements of (1) 90-104 mm, (2) 105-119 mm, and (3) 120+ mm. For management of the St. Matthew Island blue king crab (SMBKC) fishery, 120 mm CL is used as the proxy value for the legal measurement of 5.5 mm in carapace width (CW), whereas 105 mm CL is the management proxy for mature-male size (5 AAC 34.917 (d)). Accordingly, within the model only stage-3 crab are retained in the directed fishery, and stage-2 and stage-3 crab together comprise the collection of mature males. Some justification for the 105 mm value is presented in Pengilly and Schmidt (1995), who used it in developing the current regulatory SMBKC harvest strategy. The term “recruit” here designates recruits to the model, i.e., annual new stage-1 crab, rather than recruits to the fishery. The following description of model structure reflects the Gmacs base model configuration.

## 2. Model Population Dynamics

Within the model, the beginning of the crab year is assumed contemporaneous with the NMFS trawl survey, nominally assigned a date of 1 July. Although the timing of the fishery is different each year, MMB is measured 15 February, which is the reference date for calculation of federal management biomass quantities. To accommodate this, each model year is split into `r M[[2]]$nseason` seasons ($t$) and a proportion of the natural mortality ($\tau_t$) is applied in each of these seasons where $\sum_{t=1}^{t=5} \tau_t = 1$. Each model year consists of the following processes:
\begin{enumerate}
    \item Season 1
    \begin{itemize}
        \item Beginning of the SMBKC fishing year (1 July)
        \item $\tau_1 = 0$
        \item Surveys
    \end{itemize}
    \item Season 2
    \begin{itemize}
        \item $\tau_2$ ranges from 0.05 to 0.44 depending on the time of year the fishery begins each year (i.e. a higher value indicates the fishery begins later in the year; see Table \ref{tab:smbkc_fishery})
    \end{itemize}
    \item Season 3
    \begin{itemize}
        \item $\tau_3 = 0$
        \item Fishing mortality applied
    \end{itemize}
    \item Season 4
    \begin{itemize}
        \item $\tau_4 = 0.63 - \sum_{i=1}^{i=4} \tau_i$
        \item Calculate MMB (15 February)
    \end{itemize}
    \item Season 5
    \begin{itemize}
        \item $\tau_5 = 0.37$
        \item Growth and molting
        \item Recruitment (all to stage-1)
    \end{itemize}
\end{enumerate}
The proportion of natural mortality ($\tau_t$) applied during each season in the model is provided in Table \ref{tab:m_prop}. The beginning of the year (1 July) to the date that MMB is measured (15 February) is 63% of the year. Therefore 63% of the natural mortality must be applied before the MMB is calculated. Because the timing of the fishery is different each year $\tau_2$ is different each year and thus $\tau_4$ differs each year.

```{r m_prop, results = "asis"}
# A <- M[[2]]
# df <- data.frame(A$mod_yrs, A$m_prop)
# names(df) <- c("Year","Season 1","Season 2","Season 3","Season 4","Season 5")
# tab <- xtable(df, digits = c(0,0,2,2,2,2,2), caption = "Proportion of the natural mortality ($\\tau_t$) that is applied during each season ($t$) in the model.", label = "tab:m_prop")
# print(tab, caption.placement = "top", include.rownames = FALSE, sanitize.text.function=function(x){x})
```

With boldface lower-case letters indicating vector quantities we designate the vector of stage abundances during season $t$ and year $y$ as
\begin{equation}
    \boldsymbol{n}_{t,y} = n_{l,t,y} = \left[ n_{1,t,y}, n_{2,t,y}, n_{3,t,y} \right]^\top.
\end{equation}
The number of new crab, or recruits, of each stage entering the model each season $t$ and year $y$ is represented as the vector $\boldsymbol{r}_{t,y}$. The SMBKC formulation of Gmacs specifies recruitment to stage-1 only during season $t=5$, thus the recruitment size distribution is
\begin{equation}
    \phi_l = \left[ 1, 0, 0 \right]^\top,
\end{equation}
and the recruitment is
\begin{equation}
  \boldsymbol{r}_{t,y} = 
  \begin{cases}
    0 &\text{for} \quad t<5\\
    \bar{R} \phi_l \delta^R_y &\text{for} \quad t=5.
  \end{cases}
\end{equation}
where $\bar{R}$ is the average annual recruitment and $\delta^R_y$ are the recruitment deviations each year $y$
\begin{equation}
    \delta^R_y \sim \mathcal{N} \left( 0, \sigma_R^2 \right).
\end{equation}
Using boldface upper-case letters to indicate a matrix, we describe the size transition matrix $\boldsymbol{G}$ as
\begin{equation}
  \boldsymbol{G} = \left[ \begin{array}{ccc}
    1 - \pi_{12} - \pi_{13} & \pi_{12} & \pi_{13} \\
    0 & 1 - \pi_{23} & \pi_{23} \\
    0 & 0 & 1 \end{array} \right],
\end{equation}
with $\pi_{jk}$ equal to the proportion of stage-$j$ crab that molt and grow into stage-$k$ within a season or year. 

The natural mortality each season $t$ and year $y$ is
\begin{equation}
    M_{t,y} = \bar{M} \tau_t + \delta_y^M \text{ where } \delta_y^M \sim \mathcal{N} \left( 0, \sigma_M^2 \right)
\end{equation}
Fishing mortality by year $y$ and season $t$ is denoted $F_{t,y}$ and calculated as
\begin{equation}
    F_{t,y} = F_{t,y}^\text{df} + F_{t,y}^\text{tb} + F_{t,y}^\text{fb}
\end{equation}
where $F_{t,y}^\text{df}$ is the fishing mortality associated with the directed fishery, $F_{t,y}^\text{tb}$ is the fishing mortality associated with the trawl bycatch fishery, $F_{t,y}^\text{fb}$ is the fishing mortality associated with the fixed bycatch fishery. Each of these are derived as
\begin{align}
    F_{t,y}^\text{df} &= \bar{F}^\text{df} + \delta^\text{df}_{t,y} \quad \text{where} \quad \delta^\text{df}_{t,y} \sim \mathcal{N} \left( 0, \sigma^2_\text{df} \right), \notag\\
    F_{t,y}^\text{tb} &= \bar{F}^\text{tb} + \delta^\text{tb}_{t,y} \quad \text{where} \quad \delta^\text{df}_{t,y} \sim \mathcal{N} \left( 0, \sigma^2_\text{tb} \right), \notag\\
    F_{t,y}^\text{fb} &= \bar{F}^\text{fb} + \delta^\text{fb}_{t,y} \quad \text{where} \quad \delta^\text{df}_{t,y} \sim \mathcal{N} \left( 0, \sigma^2_\text{fb} \right),
\end{align}
where $\delta^\text{df}_{t,y}$, $\delta^\text{tb}_{t,y}$, and $\delta^\text{fb}_{t,y}$ are the fishing mortality deviations for each of the fisheries, each season $t$ during each year $y$, $\bar{F}^\text{df}$, $\bar{F}^\text{tb}$, and $\bar{F}^\text{fb}$ are the average fishing mortalities for each fishery. The total mortality $Z_{l,t,y}$ represents the combination of natural mortality $M_{t,y}$ and fishing mortality $F_{t,y}$ during season $t$ and year $y$
\begin{equation}
    \boldsymbol{Z}_{t,y} = Z_{l,t,y} = M_{t,y} + F_{t,y}.
\end{equation}
The survival matrix $\boldsymbol{S}_{t,y}$ during season $t$ and year $y$ is
\begin{equation}
  \boldsymbol{S}_{t,y} = \left[ \begin{array}{ccc}
    1-e^{-Z_{1,t,y}} & 0 & 0 \\
    0 & 1-e^{-Z_{2,t,y}} & 0 \\
    0 & 0 & 1-e^{-Z_{3,t,y}} \end{array} \right].
\end{equation}

The basic population dynamics underlying Gmacs can thus be described as
\begin{align}
    \boldsymbol{n}_{t+1,y} &= \boldsymbol{S}_{t,y} \boldsymbol{n}_{t,y}, &\text{ if } t<5 \notag\\
    \boldsymbol{n}_{t,y+1} &= \boldsymbol{G} \boldsymbol{S}_{t,y} \boldsymbol{n}_{t,y} + \boldsymbol{r}_{t,y} &\text{ if } t=5.
\end{align}


## 3. Model Data

Data inputs used in model estimation are listed in Table \ref{tab:model_data}.
\begin{table}[ht]
\centering
\caption{Data inputs used in model estimation.} 
\label{tab:model_data}
\begin{tabular}{lll}
  \hline
  Data & Years & Source \\
  \hline
  Directed pot-fishery retained-catch number & 1978/79 - 1998/99 & Fish tickets \\
  (not biomass) & 2009/10 - 2015/16 & (fishery closed 1999/00 - 2008/09)\\
  \hline
  Groundfish trawl bycatch biomass & 1992/93 - 2015/16 & NMFS groundfish observer program \\
  \hline
  Groundfish fixed-gear bycatch biomass & 1992/93 - 2015/16 & NMFS groundfish observer program \\
  \hline
  NMFS trawl-survey biomass index & & \\
  (area-swept estimate) and CV & 1978-2016 & NMFS EBS trawl survey \\
  \hline
  ADF\&G pot-survey abundance index & & \\
  (CPUE) and CV & Triennial 1995-2016 & ADF\&G SMBKC pot survey \\
  \hline
  NMFS trawl-survey stage proportions & & \\
  and total number of measured crab & 1978-2016 & NMFS EBS trawl survey \\
  \hline
  ADF\&G pot-survey stage proportions & & \\
  and total number of measured crab & Triennial 1995-2016 & ADF\&G SMBKC pot survey \\
  \hline
  Directed pot-fishery stage proportions & 1990/91 - 1998/99 & ADF\&G crab observer program \\
  and total number of measured crab & 2009/10 - 2015/16 & (fishery closed 1999/00 - 2008/09) \\
  \hline
\end{tabular}
\end{table}


## 4. Model Parameters

Table \ref{tab:fixed_pars} lists fixed (externally determined) parameters used in model computations. In all scenarios, the stage-transition matrix is
\begin{equation}
  \label{eq:size_transition}
  \boldsymbol{G} =
  \left[ \begin{array}{ccc}
    0.2 & 0.7 & 0.1 \\
    0 & 0.4 & 0.6 \\
    0 & 0 & 1 \end{array} \right]
\end{equation}
which is the combination of the growth matrix and molting probabilities.
\begin{table}[ht]
\centering
\caption{Fixed model parameters for all scenarios.} 
\label{tab:fixed_pars}
\begin{tabular}{lccl}
  \hline
  Parameter & Symbol & Value & Source/rationale \\
  \hline
  Trawl-survey catchability & $q$ & 1.0 & Default \\
  Natural mortality & $M$ & 0.18 $\text{yr}^{-1}$ & NPFMC (2007) \\
  Size transition matrix & $\boldsymbol{G}$ & Equation \ref{eq:size_transition} & Otto and Cummiskey (1990) \\
  Stage-1 and stage-2 & $w_{1}$, $w_{2}$ & 0.7, 1.2 kg & Length-weight equation (B. Foy, NMFS) \\
  mean weights & & & applied to stage midpoints \\
  Stage-3 mean weight & $w_{3,y}$ & Depends on year & Fishery reported average retained weight \\
  & & Table \ref{tab:length_weight} & from fish tickets, or its average, and \\
  & & & mean weights of legal males \\
  Recruitment SD & $\sigma_R$ & 1.2 & High value \\
  Natural mortality SD & $\sigma_M$ & 10.0 & High value (basically free parameter) \\
  Directed fishery & & 0.2 & 2010 Crab SAFE \\
  handling mortality & & & \\
  Groundfish trawl & & 0.8 & 2010 Crab SAFE \\
  handling mortality & & & \\
  Groundfish fixed-gear & & 0.5 & 2010 Crab SAFE \\
  handling mortality & & & \\
  \hline
\end{tabular}
\end{table}

Estimated parameters are listed in Table \ref{tab:bounds_pars} and include an estimated natural mortality deviation parameter in 1998/99 ($\delta^M_{1998}$) assuming an anomalous mortality event in that year, as hypothesized by Zheng and Kruse (2002), with natural mortality otherwise fixed at 0.18 $\text{yr}^{-1}$.
\begin{table}[ht]
\centering
\caption{The lower bound (LB), upper bound (UB), initial value, prior, and estimation phase for each estimated model parameter.} 
\label{tab:bounds_pars}
\begin{tabular}{lrrrlr}
  \hline
  Parameter & LB & Initial value & UB & Prior & Phase \\ 
  \hline
  Average recruitment $\log (\bar{R})$ & -7 & 10.0 & 20 & Uniform(-7,20) & 1 \\ 
  Stage-1 initial numbers $\log (n^0_1)$ & 5 & 14.5 & 20 & Uniform(5,20) & 1 \\ 
  Stage-2 initial numbers $\log (n^0_2)$ & 5 & 14.0 & 20 & Uniform(5,20) & 1 \\ 
  Stage-3 initial numbers $\log (n^0_3)$ & 5 & 13.5 & 20 & Uniform(5,20) & 1 \\ 
  ADF\&G pot survey catchability $q$ & 0 & 4.0 & 5 & Uniform(0,5) & 1 \\ 
  Stage-1 directed fishery selectivity 1978-2008 & 0 & 0.4 & 1 & Uniform(0,1) & 3 \\ 
  Stage-2 directed fishery selectivity 1978-2008 & 0 & 0.7 & 1 & Uniform(0,1) & 3 \\ 
  Stage-1 directed fishery selectivity 2009-2015 & 0 & 0.4 & 1 & Uniform(0,1) & 3 \\ 
  Stage-2 directed fishery selectivity 2009-2015 & 0 & 0.7 & 1 & Uniform(0,1) & 3 \\ 
  Stage-1 NMFS trawl survey selectivity & 0 & 0.4 & 1 & Uniform(0,1) & 4 \\ 
  Stage-2 NMFS trawl survey selectivity & 0 & 0.7 & 1 & Uniform(0,1) & 4 \\ 
  Stage-1 ADF\&G pot survey selectivity & 0 & 0.4 & 1 & Uniform(0,1) & 4 \\ 
  Stage-2 ADF\&G pot survey selectivity & 0 & 0.7 & 1 & Uniform(0,1) & 4 \\ 
  Natural mortality deviation during 1998 $\delta^M_{1998}$ & -3 & 0.0 & 3 & Normal(0, $\sigma^2_M$) & 4 \\ 
  Recruitment deviations $\delta^R_y$ & -7 & 0.0 & 7 & Normal(0, $\sigma_R^2$) & 3 \\ 
  Average directed fishery fishing mortality $\bar{F}^\text{df}$                 & -  & 0.2 & - & - & 1 \\
  Average trawl bycatch fishing mortality $\bar{F}^\text{tb}$                 & -  & 0.001 & - & - & 1 \\
  Average fixed gear bycatch fishing mortality $\bar{F}^\text{fb}$                 & -  & 0.001 & - & - & 1 \\
  \hline
\end{tabular}
\end{table}


## 5. Model Objective Function and Weighting Scheme

The objective function consists of the sum of several "negative log-likelihood" terms characterizing the hypothesized error structure of the principal data inputs (Table \ref{tab:likelihood_components}). A lognormal distribution is assumed to characterize the catch data and is modelled as
\begin{align}
  \sigma_{t,y}^\text{catch} &= \sqrt{\log \left( 1 + \left( \mathit{CV}_{t,y}^\text{catch} \right)^2 \right)}\\
  \delta_{t,y}^\text{catch} &= \mathcal{N} \left( 0, \left( \sigma_{t,y}^\text{catch} \right)^2 \right)
\end{align}
where $\delta_{t,y}^\text{catch}$ is the residual catch. The relative abundance data is also assumed to be lognormally distributed
\begin{align}
  \sigma_{t,y}^\text{I} &= \frac{1}{\lambda} \sqrt{\log \left( 1 + \left( \mathit{CV}_{t,y}^\text{I} \right)^2 \right)}\\
  \delta_{t,y}^\text{I} &= \log \left( I^\text{obs} / I^\text{pred} \right) / \sigma_{t,y}^\text{I} + 0.5 \sigma_{t,y}^\text{I}
\end{align}
and the likelihood is
\begin{equation}
  \sum \log \left( \delta_{t,y}^\text{I} \right) + \sum 0.5 \left( \sigma_{t,y}^\text{I} \right)^2
\end{equation}

Gmacs calculates standard deviation of the normalised residual (SDNR) values
and median of the absolute residual (MAR) values for all abundance indices and
size compositions to help the user come up with resonable likelihood weights.
For an abundance data set to be well fitted, the SDNR should not be much
greater than 1 (a value much less than 1, which means that the data set is
fitted better than was expected, is not a cause for concern). What is meant by
"much greater than 1" depends on $m$ (the number of years in the data set).
Francis (2011) suggests upper limits of 1.54, 1.37, and 1.26 for $m$ = 5, 10,
and 20, respectively. Although an SDNR not much greater than 1 is a necessary
condition for a good fit, it is not sufficient. It is important to plot the
observed and expected abundances to ensure that the fit is good.

Gmacs also calculates Francis weights for each of the size composition data
sets supplied (Francis 2011). If the user wishes to use the Francis iterative
re-weighting method, first the weights applied to the abundance indices should
be adjusted by trial and error until the SDNR (and/or MAR) are adequte. Then
the Francis weights supplied by Gmacs should be used as the new likelihood
weights for each of the size composition data sets the next time the model is
run. The user can then iteratively adjust the abundance index and size
composition weights until adequate SDNR (and/or MAR) values are achieved,
given the Francis weights.


## 6. Estimation

The model was implemented using the software AD Model Builder (Fournier et al.
2012), with parameter estimation by minimization of the model objective
function using automatic differentiation. Parameter estimates and standard
deviations provided in this document are AD Model Builder reported values
assuming maximum likelihood theory asymptotics.
